### 2019-03-12 proposals/TESS_GI_cycle02/gully_mcd_2016_prop.tex
First commit to the TESS GI Cycle 2 proposal

\documentclass[11pt,aas_macros]{article}
\usepackage{graphicx}    
\oddsidemargin 0pt
\textwidth 6.5in
\topmargin -0.65in
\textheight 9.25in

\usepackage{aas_macros}
\usepackage[comma,authoryear]{natbib}
\bibpunct{(}{)}{;}{a}{}{;}

\usepackage{caption}
\usepackage{subcaption}

\usepackage[backref,breaklinks,colorlinks,urlcolor=blue,citecolor=blue,linkcolor=blue]{hyperref}
\bibliographystyle{plainnat}

% This simple template will produce text files which conform with the
% McDonald Observatory format requirements.  
%
% This template can be used for a single section by just uncommenting
% the header for that section or for all needed sections in one 
% document by uncommenting all the headers and the newpage commands.
% The text is then entered in that section.
%
% The science justification is limited to 1 page of text plus 1 page
% of figures and/or references (the user must control this limit).
% The results from previous observing time is also limited to 1 page.
% The other sections are not limited in length but should not
% be used to add science justification.

\input{mgs_setup}

\begin{document}

\begin{center}{\bf Science Justification}\end{center}

Stars form by direct gravitational collapse of molecular cloud cores.  By $\sim$1 Myr the envelopes of gas and dust have usually accreted onto the star, classical \emph{T-Tauri} stars).  By $\sim$3-10 Myr the circumstellar disks are usually gone (Class III, weak-lined \emph{T-Tauri} stars).  At some time in this process, gas giant planets form.  Observations of young stars between 0 and 10 Myr anchor the overall ``cartoon picture'' understanding of star and planet formation.

The understanding of the process is incomplete.  Measurements for intrinsic stellar properties---luminosity, effective temperature, surface gravity---are known only coarsely.  Their uncertainties flow down to uncertainties in the fundamental properties mass and age, which are estimated from positions on pre main-sequence HR diagrams (\emph{c.f} Figure \ref{fig:gull}).

For a single presumably coeval stellar cluster, the pre main-sequence HR diagrams exhibit large ($\delta t_{\ast} \sim t_{\ast}$) apparent spreads in age \citep[e.g.][]{2011A&A...534A..83R}. The large spreads in these pre main-sequence HR diagrams is controversial: some take it as evidence for intrinsic age spreads, some take it as evidence for non static accretion history \citep{2009ApJ...702L..27B, 2010ARA&A..48..581S}.

Still others take the large apparent age spreads in pre main-sequence HR diagrams as evidence for the limited physics included in the pre main-sequence evolutionary model isochrones.  Most modern evolutionary models do not yet include the effects of starspots or magnetic fields, despite observations betraying their presence, \emph{e.g.} optical monitoring and Zeeman doppler imaging \citep{2008A&A...479..827G,2014MNRAS.444.3220D}.

Sunspots are probably responsible for biases in stellar effective temperatures derived by different methods.  For example, the APOGEE spectrograph ($1.5-1.70 \;\mu$m at $R=22,500$) measured effective temperatures for 3493 young stars finding offsets in $\teff$ of typically 200$-$500 K and as high as 1000 K compared to previous studies \citep{2014ApJ...794..125C}.  Low resolution optical spectroscopy shows a typical spread of 200 K in the spectral type-to-effective-temperature conversion scale \citep{2014ApJ...786...97H}.

The importance of starspots needs to be empirically evaluated to assess systematic biases of pre-main sequence stellar evolutionary models.

In this proposal, we aim to directly measure the effect of starspots on derived stellar parameters.  Starspot characterization was identified as the most important direction for evolutionary model improvement\footnote{alongside magnetic field characterization} at the recent IAU Symposium 314, ``Young Stars \& Planets Near the Sun'' (for example, see \citet{2015ApJ...807..174S} and \citet{2015arXiv150706460F}).  Specifically, stellar evolution models including the effect of starspots can make a coeval 10 Myr population exhibit apparent age spreads of 3$-$10 Myr, with derived masses biased towards lower masses.  The observations of sunspots are lagging behind the theory.  In the Section ``Results from Previous Observing Time at McDonald'', we demonstrate that IGRINS is uniquely suited to constrain the areal coverage fraction and temperature of starspots.  The proposed work will expand the sample from 1 source to $N\sim10$ sources.

The proposed research will address a key question posed by the Astro2010 Decadal Review Panel on Planetary Systems and Star Formation: ``What Is the Origin of the Stellar Mass Function?''.  Specifically, this work will directly address the bullet point ``More investment in theoretical pre-main sequence evolutionary tracks and their calibration''.

The proposed research has notable side benefits.  In the first life cycle of the data we will focus on starspots, but not magnetic fields, due to the enhanced analysis complexity.  The data can be repurposed to measure magnetic fields through Zeeman broadening.  The spectra of diskless young stars will also serve as perfect comparison sources for ongoing research to measure disk excess emission through photospheric line veiling.  Other discovery space will be available, since our sources have abundant auxiliary data from \emph{K2} cycle 2 and \emph{Spitzer}, among others.

\newpage
\begin{center}{\bf Description of Observations \& Justification of Exposure Times}
\end{center}

We propose to acquire IGRINS observations of 15 Class III young stellar objects towards \emph{Ophiuchus} and \emph{Upper Scorpius} selected based on their \emph{K2} cycle 2 variability and \emph{lack of} disk excess emission from WISE.  These are all point sources, and should be observed in a typical ABBA nod pattern to facilitate sky subtraction.  The Table below lists the target EPIC Identifier, 2MASS Identifier, Peak-to-valley light curve amplitude variation, an estimated rotation period in days, H-band magnitude, and estimated total exposure time in minutes.  \\

\begin{tabular}{cccccc}
\hline
& & \multicolumn{2}{c}{\emph{K2} data} \\
\cline{3-4}
EPIC ID & 2MASS ID & LC Amp. & Period & $H$ & Exp. time \\
  &   & P$-$V \% & \emph{d} & \emph{mag} & \emph{m} \\
\hline
204876697 & 2MASS J16080141$-$2027416 & 20  &  	9	& 9.5	 &   20  \\
203785905 & 2MASS J16281385$-$2456113 & 15  &  	10	& 10.3   & 	20  \\
203908611 & 2MASS J16241586$-$2427352 & 15  &  	2.5	& 9.6	 &   20  \\
204264641 & 2MASS J16140211$-$2301021 & 20  &  	3	& 8.8	 &   2  \\
204566404 & 2MASS J16354836$-$2148396 & 30  &  	2	& 8.6	 &   2  \\
205034491 & 2MASS J16115626$-$1943229 & 15  &  	7	& 10.1   & 	20  \\
205154017 & 2MASS J16064385$-$1908056 & 20  &  	7	& 9.4	 &   20  \\
205164400 & 2MASS J16093378$-$1904562 & 15  &  	14	& 10.1   & 	20  \\
205483258 & 2MASS J16232454$-$1717270 & 23  &  	5	& 9.9	 &   20  \\
203440253 & 2MASS J16252883$-$2607538 &  2  &   7	& 10.1   & 	20  \\
203710077 & 2MASS J15554883$-$2512240 &  2  &   4	& 8.4	 &   2  \\
204330922 & 2MASS J16273956$-$2245230 &  3  &   5	& 8.3	 &   2  \\
204519031 & 2MASS J16004056$-$2200322 &  3  &   2.5	& 8.5	 &   2  \\
204895521 & 2MASS J16064751$-$2022322 &  2  &   7	& 10.1   & 	20  \\
205117205 & 2MASS J16101473$-$1919095 &  2  &   7	& 10.3   & 	20  \\
\hline
\end{tabular}

\hfill \break

The target list was assembled in the following way.  We required little or no mid-infrared excess by selecting WISE color $W1-W3<0.9$.  We required G, K, or M spectral types, and $H-$ and $K-$ band magnitudes less than 11.  We visually inspected the K2 cycle 2 light curves of candidates, selecting sources into two groups: those with above average sinusoidal variability, and those with below average sinusoidal variability (\emph{c.f.} Figure \ref{fig:tiger}).  We defined the average variability as 5\%, which is the interquartile range of $K2$ cycle 2 light curves\footnote{We used the \citet{2014PASP..126..948V} light curves, which still contain some instrumental artifacts. So the average variability is probably slightly overestimated.} of 1658 young stellar objects or candidate young stellar objects towards \emph{Ophiuchus} or \emph{Upper Scorpius}.  Most sources have been vetted for multiplicity via non-redundant aperture masking or high contrast imaging.

The exposure time was estimated from the $K-$band magnitude scaling relation available on the IGRINS wiki, based on a desired signal-to-noise ratio $S/N \sim$ 100, assuming ABBA nod quads of 300 or 30 seconds each.  The desired $S/N$ stems from our experience fitting two-component photosphere starspot models that exhibit deviations from single component photospheres at the few percent level.  Our analysis method employs \emph{full spectrum fitting}, leveraging \emph{all $N\sim10,000$ of the IGRINS spectral channels}.  A non-detection of a two-component photosphere will still provide a useful upper limit to the relative areal coverage and temperature contrast of starspots.

The total exposure time requested is about 3.5 hours if all sources are observed with a single epoch.  We will also acquire A0V standard stars for telluric absorption correction.  These will add about 1 hour of overhead per night.  Assuming an observing efficiency of about 70\% (for target acquisition) we are requesting 7 hours of observing time.

%\newpage
\begin{center}{\bf Description \& Justification of Special Constraints}\end{center}

\emph{Ophiuchus} and \emph{Scorpius} are above 2 airmasses for over 2 hours per night in Trimester 2.  All the sources are centered around Right Ascension of about 16 hours, so we are requesting half-nights around $16^h$ local sidereal time-- early morning in April to late evening in July.

For some of the high-variability sources with favorable periods and phases, we are also requesting multi-epoch observations.  Specifically, we seek spectrum acquisition near the maximum and minimum phases in the light curve, where the effect of starspot presence or absence will be extremized.  We will leverage LCOGT and other ground-based photometric monitoring facilities to predict the phase of variability leading up to IGRINS observations.  Because of the short periods (2-10 days), these multi-epoch observations could be accomplished at the beginning and ending of, say, a 2-4 half-night observing run.  So no special scheduling constraint should be necessary.  For the most flexibility, the multi-epoch observations could be taken in an unofficial queue style fashion, in concert with the flexible needs of other IGRINS observers.  These multi-epoch observations could be accomplished in one additional half-night, or spread out over a week of a mini-queue mode.

\newpage
\begin{center}{\bf Results from Previous Observing Time at McDonald}\end{center}

On November 15, 2015 we acquired IGRINS miniqueue observations of LkCa4, a weak-lined \emph{T-Tauri} star.  We targeted this source for its likelihood of large areal coverage of starspots, as evinced by large-amplitude photometric variability  detected in $BVRI$ bands.  Its period is 3.37$\pm$0.01 days \citep{1993AJ....106.1608V,1994IBVS.4042....1G}.  Recent optical spectro-polarimetry showed evidence for hot or cool starspots covering an estimated $25\%$ of the stellar surface \citep{2014MNRAS.444.3220D}.  The source has little or no accretion ($\log{\dot M} < 8.1$), no detection of mid-infrared nor mm excess, and no stellar companion, so its spectrum should be devoid of complicating factors like near-IR excess veiling and accretion excess.  This source is the ideal candidate for direct measurement of two-temperature photosphere due to starspots.

We combined our IGRINS spectroscopy with existing ESPaDOnS panchromatic optical echelle spectra to provide 88 spectral orders between 5100 and 25000 \AA.  Our analysis procedure involves fitting the entire spectrum with pre-computed stellar model grids \citep{2013A&A...553A...6H}, and employing a new flexible likelihood function for spectroscopic inference \citep{2015ApJ...812..128C}.  The existing implementation \emph{simultaneously} fits both the intrinsic stellar parameters, $\vt_{\ast} = \{\teff, \logg, \Z \}$, and extrinsic stellar parameters, $\vt_{\rm ext} = \{\sigma_v, \vsini, v_r, \Omega, A_V\}$, with all $N_{\mathrm{ord}}$ spectral orders sharing the same stellar parameters\footnote{The implementation \emph{also} simultaneously fits a small-amplitude polynomial to each spectral order to adjust for slight imperferfections in the continuum shape left over from the blaze correction in the IGRINS pipeline.  The coefficients of the polynomials are treated as nuisance parameters.  This strategy makes us resilient against the pitfalls of pseudo-continuum-fitting, and effortlessly propagates the uncertainty attributable to the continuum level into the uncertainty in the astrophysically interesting stellar parameters.}.  We modified the default implementation by chunking the spectrum into 58 of its cleanest spectral orders and deriving the stellar parameters $\vt_{\ast,m}$ \emph{independently} in each $m^{\mathrm{th}}$ order.  We therefore have 58 unique estimates for $\teff$ as a function of wavelength, which we compile in Figure \ref{fig:teffOrder}.  The figure shows a dramatic effect: \textbf{The derived effective temperature is on-average about 800 K cooler in the $K-$band than in the optical.}

We interpret the cooler temperatures derived at $K-$band as the presence of sunspots contributing more flux at longer wavelengths, as shown in the bottom panel of Figure \ref{fig:teffOrder}.  Here we show the absolute flux ratio, $f_{\lambda, B} / f_{\lambda, A}$ at two different temperatures: $T_\textrm{eff,A} = 4100$ and $T_\textrm{eff,B} = 3300$.  We present the ratio of two synthetic spectra from Phoenix (blue solid line), and the black body ratio (red dashed line).

We further performed a plausibility check to constrain the effect size of starspots.  We generated flux calibrated spectra at two temperatures, and forward-modeled them to resemble the LkCa4 IGRINS and ESPaDOnS spectra in all other ways (\emph{i.e.} $\logg, \Z, \vt_{\rm ext}$), including noise and calibration parameters.  We coadded the spectra in a mixture model:  $ \mathsf{M}_{mix} = c \cdot \mathsf{M}_A(T_\textrm{eff,A}) + (1-c) \cdot \mathsf{M}_B(T_\textrm{eff,B})$.  We chose a fill factor of starspots of 30\%.  We then re-ran our single-temperature fitting procedure on the two-temperature, synthetic, noised-up data to see what stellar parameters one would na\"{\i}vely derive.  The noised-up optical data yields $\teff \sim 4100$ K, close to that of the A component.  We are awaiting the results from $K-$ band at the time of writing.

Our ongoing work is focused on several improvements to the whole-spectrum fitting procedure.  First we are adding support for directly fitting a mixture model, $\mathsf{M}_{mix}$.  This will provide a \emph{measurement} of the star spot temperature and areal coverage fraction.  Mixture model fitting capability will enable true whole-spectrum fitting, a more powerful on-label use of the fitting machinery.

We are also introducing improvements in residual spectrum outlier rejection, which will reduce some scatter in our measurements.

\begin{figure}[b]
    \centering
    \begin{subfigure}[b]{0.48\textwidth}
        \includegraphics[bb = 50 363 550 720,clip=true, width=1.0\textwidth]{figures/lkca4_hrdiag.pdf}
        \caption{A pre-main-sequence HR diagram}
        \label{fig:gull}
    \end{subfigure}
    ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
      %(or a blank line to force the subfigure onto a new line)
    \begin{subfigure}[b]{0.48\textwidth}
        \includegraphics[width=1.0\textwidth]{figures/K2_selection.pdf} 
        \caption{K2 Cycle 2 YSOs in OPH or SCO}
        \label{fig:tiger}
    \end{subfigure}
    \caption{The position of LkCa4 using effective temperatures derived from the optical or $H-$band ($\teff\sim4100$), versus the $K-$band ($\teff \sim 3300$). \emph{(b)} The green contours show the variability locus of 1658 young sources.  The black star is the position of LkCa4.  We selected sources with above- and below- average amplitude.} \label{fig:science}
\end{figure}

\begin{figure}[b]
	\centering
	\includegraphics[width=0.85\textwidth]{figures/teff_v_wl_example} 
	\caption{\emph{Top panel:} Effective temperature as derived by different spectral orders.  The pattern is consistent with more than 25\% of the stellar surface being covered in cool spots.  \emph{Bottom panel:}  The ratio of flux for a patch of stellar photosphere with $\teff=3300$ K to $\teff=4100$ K. }
	\label{fig:teffOrder}
\end{figure}

\bibliography{gully_mcd_2016_prop}

\end{document}

### 2019-03-12 proposals/TESS_GI_cycle02/tessgi_gully_cycle2.tex
First commit to the TESS GI Cycle 2 proposal

%% LaTeX template for the science justification & technical
%% feasibility to be submitted as part of a TESS Guest Investigator
%% Program proposal. This template is based on the proposal template
%% used by the NuSTAR mission.
%%
%% TESS Guest Investigator Proposal Cycle 2 template
%% V1.0
%% 2017-08-04
%% V1.1
%% 2019-02-07

%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%% DOCUMENT FORMAT %%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% The default font was chosen to be easily readable while allowing
%% sufficient material to be included.

%% Please note that the proposal will be printed on US Letter size paper,
%% 8.5 in x 11 in, and that formatting the text for other sizes will
%% generally cause layout problems and may result in text being cut
%% off near the edges. PLEASE DO NOT CHANGE THE 'LETTERPAPER' OPTION
%% IN THE DOCUMENTCLASS COMMAND.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%% Default format: 11pt single column %%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% NOTE: NASA ROSES requires body font size to be no smaller than 15
%% characters per inch (equivalent to Times Roman 12 point).
%%
%% Minimum margin size is 1 inch from top, bottom, and sides.

\documentclass[letterpaper,11pt]{article}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%% HOW TO INCLUDE FIGURES %%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% Please see the ``Included packages'' section below.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%% Included packages %%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\usepackage{graphics,graphicx}
\usepackage{aas_macros}
\usepackage[comma,authoryear]{natbib}
\bibpunct{(}{)}{;}{a}{}{;}

\usepackage{caption}
\usepackage{subcaption}

\usepackage[backref,breaklinks,colorlinks,urlcolor=blue,citecolor=blue,linkcolor=blue]{hyperref}
\bibliographystyle{plainnat}
%\usepackage{cleveref}
\input{mgs_setup}

%% Feel free to modify the included packages list to use your
%% favorite packages.

%% In the graphics and graphicx packages, Postscript and eps figures
%% can be included using the \includegraphics command. The graphics
%% package is part of standard LaTeX2e and provides a basic way of including a
%% figure. The graphicx package is not standard, but extends the
%% \includegraphics command to make it more user-friendly. If graphicx
%% is not available on your system please remove it from the list of
%% included packages above.

%% Syntax:
%% In the graphics package:
%%
%% \begin{figure}
%% \includegraphics[llx,lly][urx,ury]{file}
%% \end{figure}
%%
%% where ll denotes 'lower left' and ur 'upper right' and the x and y
%% values are the coordinates of the PostScript bounding box in
%% points. There are 72 points in an inch.
%%
%% In the graphicx package:
%%
%% \begin{figure}
%% \includegraphics[key=val,key=val,...]{file}
%% \end{figure}
%%
%% where some of the useful keys are: angle, width, height,
%% keepaspectratio (='true' or 'false') and scale. Bounding box values
%% can be given as [bb=llx lly urx ury].
%%
%% In either case you have to use LaTeX figure placement commands to
%% position the figure on the page; \includegraphics will not do
%% that. Both these commands also have other options that are listed
%% in the LaTeX manual (for the graphics package) and in 'The LaTeX
%% Graphics Companion' (for the graphicx package).

%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%% Page dimensions %%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%

\setlength{\textwidth}{6.5in}
\setlength{\textheight}{9in}
\setlength{\topmargin}{-0.0625in}
\setlength{\oddsidemargin}{0in}
\setlength{\evensidemargin}{0in}
\setlength{\headheight}{0in}
\setlength{\headsep}{0in}
\setlength{\hoffset}{0in}
\setlength{\voffset}{0in}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%% Section heading format %%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\makeatletter
\renewcommand{\section}{\@startsection%
{section}{1}{0mm}{-\baselineskip}%
{0.5\baselineskip}{\normalfont\Large\bfseries}}%
\makeatother

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%% Some Useful Abbreviations %%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newcommand{\tess}{{\it TESS}}
\newcommand{\jwst}{{\it JWST}}
\newcommand{\kepler}{{\it Kepler}}
\newcommand{\ktwo}{{K2}}
\newcommand{\hst}{{\it HST}}
\newcommand{\msun}{$M_{\odot}$}
\newcommand{\rsun}{$R_{\odot}$}
\newcommand{\lsun}{$L_{\odot}$}
\newcommand{\re}{$R_{\oplus}$}
\newcommand{\me}{$M_{\oplus}$}
\newcommand{\rj}{$R_{\textrm{\scriptsize Jup}}$}
\newcommand{\mj}{$M_{\textrm{\scriptsize Jup}}$}
\newcommand{\ms}{m~s$^{-1}$}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%% Start of document %%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}
\pagestyle{plain}
\pagenumbering{arabic}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%% Title of proposal %%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{center}
\bfseries\uppercase{%
%%
%% ENTER TITLE OF PROPOSAL BELOW THIS LINE
REPLACE THIS LINE WITH THE TITLE
%%
%%
}
\end{center}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%% Body of science justification %%%%%
%%%%% and technical feasibility     %%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\noindent{The recommended sections for a TESS GI proposal are shown below. Feel free to change section
headings as necessary, but this is the suggested minimal information that should be included in the proposal.
This Science/Technical section of the proposal is limited to 4 pages including figures and references.}

\section{Introduction}

Summarize the problem being addressed and give an overview of how your investigation will help.
Why \tess, why now?

\section{Scientific Justification}

Provide text and figures that justify the scientific need for \tess\ observations and analyses here.
Justify your choice of cadence. When applicable, justify your choice of new 2 min cadence observations.
If your program includes theoretical, simulation, or ground-based observing components (limited to no
more than 30\% effort), describe why these efforts are critical in addition to \tess\ observations.

For example, see \cite{2015ApJ...807..174S} and \cite{2015arXiv150706460F}).  Specifically, stellar evolution models

%% Trigger Criteria section:
%% comment this section out on proposals not asking for
%% Target of Opportunity (ToO) observations.

\subsection{({\it Only} For ToO Proposals) Trigger Criteria}

If the proposed investigation includes Targets of Opportunity (ToO's), describe also the circumstances
under which a ToO is triggered, an estimated duration of the event(s), and an estimated probability for
triggering the observations. Also discuss the potential science impact imposed by the delay in upload
of the event due to \tess\ orbit/uplink constraints.

\section{Analysis Plan}

Discuss how you plan to analyze the \tess\ data. This includes the development of software tools.

\section{Technical Feasibility}

Provide text and figures showing that the proposed \tess\ investigations are feasible; consider the
\tess\ survey strategy, target observability, and required signal-to-noise, etc. The \tess\ Science
Support Center (\href{https://heasarc.gsfc.nasa.gov/docs/tess/}{TSSC}) makes several tools available to help estimate these quantities.

\section{Expected Impact}

Summarize the expected science return of the proposed investigations and the expected benefit
to the community, including new data products and software tools to be made publicly available.

\section{Work Plan}

Provide a brief (1 paragraph) work plan that provides details on how the proposed effort will be carried out, including the allocation of effort amongst investigators. Investigators who are proposing to continue a program that was selected for funding in Cycle 1 must describe why additional funds are required in Cycle 2. All proposals requesting funds must also provide upon submission a bottom-line budget number in the provided field of the ARK submission form; this number should not be included in the body of the proposal.

\section{References}

List of references. References {\it are included} when considering the proposal 4 page limit.
\bibliography{tessgi_gully_cycle2}

\section{Target Table}

When necessary to justify your proposal, provide a list of targets using the below example as
a template for format. This target table is designed to aid reviewers and need only provide a
representative sample of the complete target list uploaded to RPS. Full target tables should be
submitted electronically with the Phase-1 proposal. Please limit this target table to only 1 page.
The table is {\it not included} in the 4 page limit of the Science/Technical section. Proposals including
a target table may be up to 5 pages in length, but only the target table may appear on page 5.

\begin{center}
\begin{tabular}{ | c | c | c | c | c | c | c | }
\hline
TIC ID          &      Common      &     RA             &      Dec          &      TESS       &       Obj.        &      Comments \\
                    &      Name           &     (deg)          &      (deg)        &      mag         &       Type       &                         \\
\hline
\hline
388857263  &  Prox Cen          &  217.428793  &  -62.679592  &  7.36             &    M Dwarf    & 2 min cad., RV planet \\ \hline
                    &                           &                       &                      &                      &                     &                                     \\ \hline
                    &                           &                       &                      &                      &                     &                                     \\ \hline
                    &                           &                       &                      &                      &                     &                                     \\ \hline
                    &                           &                       &                      &                      &                     &                                     \\ \hline
\end{tabular}
\end{center}

%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%% End of document %%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%

\end{document}

### 2019-03-14 proposals/TESS_GI_cycle02/tessgi_gully_cycle2.tex
Start putting the outline into the proposal

@@
 %% ENTER TITLE OF PROPOSAL BELOW THIS LINE
-REPLACE THIS LINE WITH THE TITLE
+Starspots confound derived stellar properties and a concept for probabilistic mitigation
 %%
@@
 
-\noindent{The recommended sections for a TESS GI proposal are shown below. Feel free to change section
-headings as necessary, but this is the suggested minimal information that should be included in the proposal.
-This Science/Technical section of the proposal is limited to 4 pages including figures and references.}
-
 \section{Introduction}
@@
 
+\subsection{Starspots as confounders of isochronal ages}
 Provide text and figures that justify the scientific need for \tess\ observations and analyses here.
@@
 
-\subsection{({\it Only} For ToO Proposals) Trigger Criteria}
+\subsection{Transit light source effect}
+thinga
 
-If the proposed investigation includes Targets of Opportunity (ToO's), describe also the circumstances
-under which a ToO is triggered, an estimated duration of the event(s), and an estimated probability for
-triggering the observations. Also discuss the potential science impact imposed by the delay in upload
-of the event due to \tess\ orbit/uplink constraints.
+\subsection{Stellar surface structures as a gateway for brown dwarf cloud modelin}
+more yjumd
 
@@
 
-Discuss how you plan to analyze the \tess\ data. This includes the development of software tools.
+\subsection{Get amplitudes and periods for a large fraction of stars from TESS data}
+\subsection{Get starspot temperature and coverage fraction for a sample of these stars from spectroscopy}
+\subsection{Compute the ratio of TESS amplitude to total spot coverage}
+\subsection{Compute a regression of spot coverage to rotation rate and spectral subtype.}
 
@@
 
-Provide text and figures showing that the proposed \tess\ investigations are feasible; consider the
-\tess\ survey strategy, target observability, and required signal-to-noise, etc. The \tess\ Science
-Support Center (\href{https://heasarc.gsfc.nasa.gov/docs/tess/}{TSSC}) makes several tools available to help estimate these quantities.
+\subsection{Essentially measuring periods and amplitudes, which has been demonstrated}
+\subsection{There will be enough GKM stars}
+\subsection{The TESS precision will be adequate}
+\subsection{I have already applied the spectral technique to a wide range of young stars}
+\subsection{Availability of archival spectroscopy}
 

### 2019-03-14 proposals/TESS_GI_cycle02/tessgi_gully_cycle2.tex
Work plan scaffolding is there

@@
 \subsection{Get amplitudes and periods for a large fraction of stars from TESS data}
-\subsection{Get starspot temperature and coverage fraction for a sample of these stars from spectroscopy}
+\subsection{Compute the typical ratio of Kepler/K2 amplitudes to TESS amplitudes at given SpT/rotation rate}
+\subsection{Interpret the Tess/Kepler ratios as spot contrast differences}
+\subsection{Get starspot temperature and coverage fraction for a subsample of these stars from spectroscopy}
 \subsection{Compute the ratio of TESS amplitude to total spot coverage}
@@
 
-\subsection{Essentially measuring periods and amplitudes, which has been demonstrated}
+\subsection{Measuring periods and amplitudes, which has been demonstrated}
 \subsection{There will be enough GKM stars}
 \subsection{The TESS precision will be adequate}
-\subsection{I have already applied the spectral technique to a wide range of young stars}
+\subsection{Experience with the spectral technique on young stars}
 \subsection{Availability of archival spectroscopy}
@@
 
-Summarize the expected science return of the proposed investigations and the expected benefit
-to the community, including new data products and software tools to be made publicly available.
+\subsection{Probabilistic lookup table for spot temperature and spot coverage}
+\subsection{Direct estimate of spot coverage for selected high value targets}
+\subsection{Gateway to future investigation pathways}
+\subsubsection{Trends in spot coverage with lightcurve periodic substructure}
+\subsubsection{Trends in spot coverage with lightcurve high frequency residual substructure}
+\subsubsection{Trends in spot coverage with lightcurve high frequency residual substructure}
 
@@
 
-Provide a brief (1 paragraph) work plan that provides details on how the proposed effort will be carried out, including the allocation of effort amongst investigators.
-Investigators who are proposing to continue a program that was selected for funding in Cycle 1 must describe why additional funds are required in Cycle 2.
-All proposals requesting funds must also provide upon submission a bottom-line budget number in the provided field of the ARK submission form; this number should not be included in the body of the proposal.
+MGS will carry out the data analysis using custom scripts that build upon the heritage of the Kepler data analysis ecosystem, including \texttt{lightkurve}, \texttt{astropy}, \texttt{celerite}, \texttt{BombScargle}, and \texttt{bokeh}.
+MGS will first apply these scripts to Kepler SAP lightcurves or detrended K2 data while in his current role as support scientist at the Kepler/K2 Guest Observer Office.
+MGS will adapt and apply these scripts to TESS data---in the event of GI proposal selection---starting in Fall 2019.
+All scripts will be publicly available and developed openly on GitHub, including tutorial Jupyter Notebooks.
+MGS will produce an interactive \texttt{bokeh} tool to visualize the Fourier decomposition of a periodic lightcurve into its constituent components.
+This tool will leverage the PI's familiarity with \texttt{bokeh}, to aid students' visualization different amplitudes.
+A machine readable database will be produced that contains the following information for each star found to possess starspot-induced lightcurve modulation: 1) the period, 2) an iteratively determined mask of stellar flares and/or spurious data artifacts, 3) a vector of Chebychev polynomial coefficients to account for any long-term normalization trends on timescales comparable to the observation window, and 4) a vector of Fourier amplitudes that can reconstruct the periodic structures in the lightcurve to within the signal-to-noise ratio of the data.
 
-\section{References}
-
-List of references. References {\it are included} when considering the proposal 4 page limit.
 \bibliography{tessgi_gully_cycle2}
 
-\section{Target Table}
-
-When necessary to justify your proposal, provide a list of targets using the below example as
-a template for format. This target table is designed to aid reviewers and need only provide a
-representative sample of the complete target list uploaded to RPS. Full target tables should be
-submitted electronically with the Phase-1 proposal. Please limit this target table to only 1 page.
-The table is {\it not included} in the 4 page limit of the Science/Technical section. Proposals including
-a target table may be up to 5 pages in length, but only the target table may appear on page 5.
-
-\begin{center}
-\begin{tabular}{ | c | c | c | c | c | c | c | }
-\hline
-TIC ID          &      Common      &     RA             &      Dec          &      TESS       &       Obj.        &      Comments \\
-                    &      Name           &     (deg)          &      (deg)        &      mag         &       Type       &                         \\
-\hline
-\hline
-388857263  &  Prox Cen          &  217.428793  &  -62.679592  &  7.36             &    M Dwarf    & 2 min cad., RV planet \\ \hline
-                    &                           &                       &                      &                      &                     &                                     \\ \hline
-                    &                           &                       &                      &                      &                     &                                     \\ \hline
-                    &                           &                       &                      &                      &                     &                                     \\ \hline
-                    &                           &                       &                      &                      &                     &                                     \\ \hline
-\end{tabular}
-\end{center}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%% End of document %%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
 \end{document}

### 2019-03-14 proposals/TESS_GI_cycle02/tessgi_gully_cycle2.tex
Writing prose for the proposal

@@
 \subsection{Starspots as confounders of isochronal ages}
-Provide text and figures that justify the scientific need for \tess\ observations and analyses here.
-Justify your choice of cadence. When applicable, justify your choice of new 2 min cadence observations.
-If your program includes theoretical, simulation, or ground-based observing components (limited to no
-more than 30\% effort), describe why these efforts are critical in addition to \tess\ observations.
-
 For example, see \cite{2015ApJ...807..174S} and \cite{2015arXiv150706460F}).  Specifically, stellar evolution models
@@
 
-\subsection{Transit light source effect}
-thinga
+\subsection{Transit Light Source Effect}
+The diminution of flux arising from occultations of the stellar disk contains information about the relative sizes of the occulting bodies.
+A commonsense exoplanet heuristic dictates that the relative transit depth $D$ (in units of ppm or $\%$) equals the ratio of the bodies' projected areas ($R_p^2/R_\star^2$).
+Addtional correction factors account for limb darkening \citep{2002ApJ...580L.171M}.
+Starspots on a stellar surface can exhibit spatial segrations\footnote{Sunspots reside on the Sun's mid-latitudes, and starspots can exist near the stellar poles, as evinced in spatially resolved interferometry of nearby giant stars \citep{2016Natur.533..217R}} that yield \emph{unocculted starspots}: spots that are present on the surface but are not traversed during event during transit.
+These unocculted starspots confound the mapping of transit depth to planet radius: a planet ``looks bigger'' in a transit depth if it blocks the brighter-on-average flux of the stellar disk.
+\textbf{You cannot measure a planet radius without making some assumptions about the extent, contrast, and distribution of starspots.}
+
+The prevailing assumptions within the exoplanet community has been that starspot covering fractions are small enough that the
 
@@
 \subsection{Probabilistic lookup table for spot temperature and spot coverage}
+The probabilistic lookup table for spot temperature and filling factor will empower exoplanet researchers to constrain biases in exoplanet transit radii.
+
 \subsection{Direct estimate of spot coverage for selected high value targets}
-\subsection{Gateway to future investigation pathways}
-\subsubsection{Trends in spot coverage with lightcurve periodic substructure}
-\subsubsection{Trends in spot coverage with lightcurve high frequency residual substructure}
-\subsubsection{Trends in spot coverage with lightcurve high frequency residual substructure}
+The spectroscopic analysis component of this work will add long-term value to the TESS legacy.
+
+\subsection{Trends in spot coverage with lightcurve periodic substructure}
+The adopted Fourier analysis technique naturally collates the relative amplitudes of periodic structures, which encode some additional information about the spatial distribution of stellar surface inhomogeneities, and/or differential rotation.
+The collated harmonics will enable exploratory analyses of trends with other observables, as recently proposed in the amount of time a spot-modulated \kepler\ lightcurve exhibits single or double dips \citep{2018ApJ...863..190B}.
 
@@
 
-MGS will carry out the data analysis using custom scripts that build upon the heritage of the Kepler data analysis ecosystem, including \texttt{lightkurve}, \texttt{astropy}, \texttt{celerite}, \texttt{BombScargle}, and \texttt{bokeh}.
-MGS will first apply these scripts to Kepler SAP lightcurves or detrended K2 data while in his current role as support scientist at the Kepler/K2 Guest Observer Office.
-MGS will adapt and apply these scripts to TESS data---in the event of GI proposal selection---starting in Fall 2019.
-All scripts will be publicly available and developed openly on GitHub, including tutorial Jupyter Notebooks.
-MGS will produce an interactive \texttt{bokeh} tool to visualize the Fourier decomposition of a periodic lightcurve into its constituent components.
-This tool will leverage the PI's familiarity with \texttt{bokeh}, to aid students' visualization different amplitudes.
-A machine readable database will be produced that contains the following information for each star found to possess starspot-induced lightcurve modulation: 1) the period, 2) an iteratively determined mask of stellar flares and/or spurious data artifacts, 3) a vector of Chebychev polynomial coefficients to account for any long-term normalization trends on timescales comparable to the observation window, and 4) a vector of Fourier amplitudes that can reconstruct the periodic structures in the lightcurve to within the signal-to-noise ratio of the data.
+The Pincipal Investigator (PI) will carry out the data analysis using custom scripts that build upon the heritage of the Kepler data analysis ecosystem, including \texttt{lightkurve}, \texttt{astropy}, \texttt{celerite}, \texttt{BombScargle}, \texttt{PyTorch}, and \texttt{bokeh}.
+The PI will first apply these scripts to Kepler/K2 lightcurves while in their current role as support scientist at the Kepler/K2 Guest Observer Office.
+The PI will relocate to the University of Texas at Austin in February 2020 as a Research Fellow.
+UT Austin undergraduate and graduate students will lead the application of TESS-adapted scripts to \tess\ FFI lightcurves.
+\tess-adapted scripts will leverage \texttt{TESS-cut} and \texttt{elenor} to produce lightcurves from FFIs.
+Half of the funding will go to summer 2020 student salaries, and half will go to support the PI.
+All reproducible scripts---including Jupyter Notebook tutorials---will be publicly available and developed openly on GitHub.
+MGS will provide students with an interactive \texttt{bokeh} tool to visualize the Fourier decomposition of a periodic lightcurve into its constituent components.
+This tool will leverage the PI's familiarity with \texttt{bokeh} to aid understanding and rapid visual spotchecking of the multidimensional analysis output.
+Finally, a High Level Science Product (HLSP) will be delivered to MAST that contains the following information for each star found to possess detectable starspot-induced lightcurve modulation: 1) the period or periods, 2) an iteratively determined outlier mask of stellar flares, occultations, or spurious data artifacts, 3) a vector of Chebychev polynomial coefficients to account for any long-term normalization trends on timescales comparable to the observation window, and 4) an $N_{\mathrm{Fourier}}$ vector of Fourier harmonic amplitudes that reconstruct the periodic structures in the lightcurve to within the signal-to-noise ratio of the data.
+MGS will advise students in the writeup of the paper(s), authored and/or coauthored by students.
 

### 2019-03-14 proposals/TESS_GI_cycle02/tessgi_gully_cycle2.tex
more prose for the proposal

@@
 %% ENTER TITLE OF PROPOSAL BELOW THIS LINE
-Starspots confound derived stellar properties and a concept for probabilistic mitigation
+Starspots confound derived stellar properties and a strategy for probabilistic mitigation
 %%
@@
 
-Summarize the problem being addressed and give an overview of how your investigation will help.
-Why \tess, why now?
+Here we propose a strategy for mitigating starspot-induced exoplanet radius biases, based on a large sample of spotted stars observed in \tess\ Cycle 2.
+The approach will deliver constraints on both the typical starspot coverage fraction and the typical starspot temperature contrast, as a function of spectral type and rotation rate.
+This mapping can then serve as a ``lookup table'' for practitioners to gauge the extent to which a given source is likely to suffer from starspot-induced exoplanet radius biases.
+No such reliable lookup table presently exists for main sequence stars, owing to the inability of differential lightcurves to account for \emph{total} starspot coverage fractions \citep{2018ApJ...865..142B}.
+
+This program will leverage the \tess\ Cycle 2 overlap with the \kepler\ prime field, and includes an option for contemporaneous ground-based observations with unique instrumentation based in the Northern hemisphere.
 
@@
 
-\subsection{Starspots as confounders of isochronal ages}
-For example, see \cite{2015ApJ...807..174S} and \cite{2015arXiv150706460F}).  Specifically, stellar evolution models
-
-%% Trigger Criteria section:
-%% comment this section out on proposals not asking for
-%% Target of Opportunity (ToO) observations.
-
 \subsection{Transit Light Source Effect}
 The diminution of flux arising from occultations of the stellar disk contains information about the relative sizes of the occulting bodies.
-A commonsense exoplanet heuristic dictates that the relative transit depth $D$ (in units of ppm or $\%$) equals the ratio of the bodies' projected areas ($R_p^2/R_\star^2$).
+A commonsense exoplanet heuristic dictates that the relative transit depth $D$ equals the ratio of the bodies' projected areas ($R_p^2/R_\star^2$).
 Addtional correction factors account for limb darkening \citep{2002ApJ...580L.171M}.
-Starspots on a stellar surface can exhibit spatial segrations\footnote{Sunspots reside on the Sun's mid-latitudes, and starspots can exist near the stellar poles, as evinced in spatially resolved interferometry of nearby giant stars \citep{2016Natur.533..217R}} that yield \emph{unocculted starspots}: spots that are present on the surface but are not traversed during event during transit.
-These unocculted starspots confound the mapping of transit depth to planet radius: a planet ``looks bigger'' in a transit depth if it blocks the brighter-on-average flux of the stellar disk.
+Starspots on a stellar surface can exhibit spatial segrations %\footnote{Sunspots reside on the Sun's mid-latitudes; starspots can exist near the stellar poles, as evinced in spatially resolved interferometry of nearby giant stars \citep{2016Natur.533..217R}.}
+that yield \emph{unocculted starspots}: spots that are present on the surface but are not traversed during transits.
+These unocculted starspots confound the mapping of transit depth to planet radius: a planet ``looks bigger'' than it really is if it blocks the brighter-on-average flux of the stellar disk \citep{2018AJ....156...91M}.
 \textbf{You cannot measure a planet radius without making some assumptions about the extent, contrast, and distribution of starspots.}
 
-The prevailing assumptions within the exoplanet community has been that starspot covering fractions are small enough that the
-
-\subsection{Stellar surface structures as a gateway for brown dwarf cloud modelin}
-more yjumd
+The prevailing assumption within the exoplanet community has been that starspot covering fractions are small enough---as in Sunspots---to be ignored.
+New theoretical models \citep{2018ApJ...853..122R} and observational evidence \citep{2016MNRAS.463.2494F} indicate that starspot covering fractions are large enough to bias transit depths beyond the level of precision of \kepler\ and \tess.
+This so-called ``Transit Light Source Effect'' (TLSE) injects significant bias into derived exoplanet \emph{densities}, since the radius error propagates into volume error with a factor of 3.
+Therefore, TLSE threatens placement and interpretation of the exoplanet mass-radius relation science requirements of the \tess\ mission.
+The interpretation of \jwst\ exoplanet transmission spectroscopy is also at risk without improved knowledge of the level of the stellar photospheric contamination \citep{2019AJ....157...11W}.
 
@@
 
-\subsection{Get amplitudes and periods for a large fraction of stars from TESS data}
+\subsection{Kepler and Tess perceive different starspot contrast}
+
+Starspots emit light, which is often parameterized as \emph{spot contrast}, $c$, where 0 means the spots are non-emitting, and $c=1$ means thes spot matches the specific intensity of the ambient photosphere, vanishing altogether.
+
+The spot contrast is wavelength dependent ($c_\lambda$), and can be thought of as possessing a spot temperature, $T_{spot}$, distinct from the ambient photosphere $T_{phot}$.
+
+The analysis technique proposed here leverages the different wavelength dependence of the \kepler\ and \tess\ system throughputs to directly measure the ratio of the spot contrasts in their respective bandpasses:
+
+$$\frac{c_{TESS}}{c_{Kepler}}(\mathrm{SpT},\;P_{\mathrm{rot}})$$
+
+The spot contrast ratio---evaluated at these two pairs of bandpasses---provides a unique solution to $T_{phot}(\mathrm{SpT},\;P_{\mathrm{rot}})$.
+Operationally, integrating the specific intensity $I_\lambda(T_{phot})$ over the system throughputs produces the expected contrast ratios.
+For now we assume $I_\lambda(T_{phot})$ can be approximated as a stellar photosphere, such as the \texttt{PHOENIX} model grid.
+
+Rotationally-modulated lightcurves exhibit semi-sinusoidal undulations as starspots (or spot groups) enter and exit the projected stellar disk.
+The most-spotted projected hemisphere is directed towards the observer at the moment of the lightcurve minimum, and the least-spotted projected hemisphere at lightcurve maximum.
+The amplitude of the lightcurve therefore equals the difference in starspot coverage fractions between the most-and-least spotted projected hemispheres \emph{multiplied by the spot contrast}.
+
+\subsection{Different amplitudes and periods for a large fraction of stars from TESS data}
+
+Rapid rotators tend to exhibit larger amplitudes of lightcurve undulations than slower rotators.
+The amplitude of the If starspots were completely non-emitting across all wavelengths, \kepler\ and \tess\ would measure the same
+
 \subsection{Compute the typical ratio of Kepler/K2 amplitudes to TESS amplitudes at given SpT/rotation rate}
@@
 
+\subsection{Stellar surface structures as a gateway for brown dwarf cloud modeling}
+
 \section{Work Plan}

### 2019-03-14 proposals/TESS_GI_cycle02/tessgi_gully_cycle2.tex
more and more prose and now more equations

@@
 \usepackage{subcaption}
+\usepackage{cancel}
 
@@
 
-\subsection{Kepler and Tess perceive different starspot contrast}
+\subsection{Kepler and TESS perceive different starspot contrast}
 
@@
 
-The spot contrast is wavelength dependent ($c_\lambda$), and can be thought of as possessing a spot temperature, $T_{spot}$, distinct from the ambient photosphere $T_{phot}$.
+The spot contrast is wavelength dependent ($c_\lambda$), and can be thought of as possessing a characteristic spot temperature, $T_{spot}<T_{phot}$ that parameterizes the wavelength dependence of the contrasts.
 
@@
 
-The spot contrast ratio---evaluated at these two pairs of bandpasses---provides a unique solution to $T_{phot}(\mathrm{SpT},\;P_{\mathrm{rot}})$.
-Operationally, integrating the specific intensity $I_\lambda(T_{phot})$ over the system throughputs produces the expected contrast ratios.
-For now we assume $I_\lambda(T_{phot})$ can be approximated as a stellar photosphere, such as the \texttt{PHOENIX} model grid.
+The spot contrast ratio---evaluated at these two pairs of bandpasses---provides a unique solution\footnote{Operationally, integrating the specific intensity $I_\lambda(T_{spot})$ over the system throughputs produces the expected contrast ratios.
+For now we assume $I_\lambda(T_{spot})$ can be approximated as a cooler stellar photosphere, such as the \texttt{PHOENIX} model grid \citep{2013A&A...553A...6H}, an inexact but adequate assumption for the large bandpasses.} to $T_{spot}(\mathrm{SpT},\;P_{\mathrm{rot}})$.
 
+\subsection{Spot-induced modulation amplitudes depend on contrast}
 Rotationally-modulated lightcurves exhibit semi-sinusoidal undulations as starspots (or spot groups) enter and exit the projected stellar disk.
 The most-spotted projected hemisphere is directed towards the observer at the moment of the lightcurve minimum, and the least-spotted projected hemisphere at lightcurve maximum.
-The amplitude of the lightcurve therefore equals the difference in starspot coverage fractions between the most-and-least spotted projected hemispheres \emph{multiplied by the spot contrast}.
+The amplitude of the lightcurve therefore equals the difference in starspot coverage fractions $\Delta f_{spot}$ between the most-and-least spotted projected hemispheres \emph{weighted by the spot contrast}:
 
-\subsection{Different amplitudes and periods for a large fraction of stars from TESS data}
+$$ A_\lambda = \Delta f_{spot} \cdot (1-c_\lambda)$$
 
-Rapid rotators tend to exhibit larger amplitudes of lightcurve undulations than slower rotators.
-The amplitude of the If starspots were completely non-emitting across all wavelengths, \kepler\ and \tess\ would measure the same
+Therefore the ratio of \tess\ to \kepler\ amplitudes yields a constraint independent of the actual spot coverage fraction:
 
-\subsection{Compute the typical ratio of Kepler/K2 amplitudes to TESS amplitudes at given SpT/rotation rate}
+$$ \frac{A_{TESS}}{A_{Kepler}} = \cancel{\frac{\Delta f_{spot}}{\Delta f_{spot}}} \cdot \frac{1-c_{TESS}}{1-c_{Kepler}} \label{eqn1}$$
+
+\subsection{Measuring Kepler/K2 and TESS amplitudes: Statistical samples}
+We see \ref{eqn1} above.
+\subsection{Measuring Kepler/K2 and TESS amplitudes: The Kepler Prime Field}
 \subsection{Interpret the Tess/Kepler ratios as spot contrast differences}
 \subsection{Get starspot temperature and coverage fraction for a subsample of these stars from spectroscopy}
-\subsection{Compute the ratio of TESS amplitude to total spot coverage}
-\subsection{Compute a regression of spot coverage to rotation rate and spectral subtype.}
+%\subsection{Compute the ratio of TESS amplitude to total spot coverage}
+%\subsection{Compute a regression of spot coverage to rotation rate and spectral subtype.}
 
@@
 \subsection{The TESS precision will be adequate}
+Rapid rotators tend to exhibit larger amplitudes of lightcurve undulations than slower rotators.
 \subsection{Experience with the spectral technique on young stars}

### 2019-03-14 proposals/TESS_GI_cycle02/tessgi_gully_cycle2.tex
More prose, and now add in a figure.

@@
 \pagenumbering{arabic}
+\setlength{\bibsep}{0.0pt}
+\let\oldbibliography\thebibliography
+\renewcommand{\thebibliography}[1]{%
+  \oldbibliography{#1}%
+  \setlength{\itemsep}{0pt}%
+}
 
@@
 
-This program will leverage the \tess\ Cycle 2 overlap with the \kepler\ prime field, and includes an option for contemporaneous ground-based observations with unique instrumentation based in the Northern hemisphere.
+This program will leverage the \tess\ Cycle 2 overlap with the \kepler\ prime field and portions of \ktwo\ fields.
 
@@
 
-Starspots emit light, which is often parameterized as \emph{spot contrast}, $c$, where 0 means the spots are non-emitting, and $c=1$ means thes spot matches the specific intensity of the ambient photosphere, vanishing altogether.
+Starspots emit light, which is often parameterized as \emph{spot contrast}, $c$, where 0 means the spots are non-emitting, and $c=1$ means thes spot matches the specific intensity of the ambient photosphere.
+The sun
 
@@
 
-The analysis technique proposed here leverages the different wavelength dependence of the \kepler\ and \tess\ system throughputs to directly measure the ratio of the spot contrasts in their respective bandpasses:
+The analysis technique proposed here leverages the different wavelength dependence of the \kepler\ and \tess\ system throughputs to directly measure the ratio of the spot contrasts in their respective bandpasses.
 
@@
 
-$$ \frac{A_{TESS}}{A_{Kepler}} = \cancel{\frac{\Delta f_{spot}}{\Delta f_{spot}}} \cdot \frac{1-c_{TESS}}{1-c_{Kepler}} \label{eqn1}$$
+$$ \frac{A_{TESS}}{A_{Kepler}} = \cancel{\frac{\Delta f_{spot}}{\Delta f_{spot}}} \cdot \frac{1-c_{TESS}}{1-c_{Kepler}} $$
 
 \subsection{Measuring Kepler/K2 and TESS amplitudes: Statistical samples}
-We see \ref{eqn1} above.
-\subsection{Measuring Kepler/K2 and TESS amplitudes: The Kepler Prime Field}
-\subsection{Interpret the Tess/Kepler ratios as spot contrast differences}
-\subsection{Get starspot temperature and coverage fraction for a subsample of these stars from spectroscopy}
-%\subsection{Compute the ratio of TESS amplitude to total spot coverage}
-%\subsection{Compute a regression of spot coverage to rotation rate and spectral subtype.}
+Ideally we would have simultaneouly observed the same targets with \kepler\ and \tess\ to measure the contrast for individual stars\footnote{Technically K2 Campaign 19 contemporaneously observed a subset of the same stars as \tess, but yielded few sources relevant to this discussion.}.
+Since this observational strategy was not possible, we can construct statistical subsamples of stars that are identical in every way, except for their observation band.
+
+The main two factors that affect spot-modulation lightcurve amplitudes are effective temperature (\emph{i.e.} mass), and rotation period (\emph{i.e.} stellar activity/magnetic fields) \citep{2014ApJS..211...24M}.
+Other important albeit unknwon factors like stellar inclination angle and multiplicity are statistically identical between the \tess\ and \kepler\ samples.
+We therefore propose to follow \citet{2014ApJS..211...24M} by constructing plots of amplitude versus period, subdivided into bins of effective temperature.
+The two samples should look identical in these plots, but the locus of \tess\ sources are expected to fall below the \kepler\ sample, as in Figure \ref{fig:predictAmps}.
+
+\begin{figure}
+\includegraphics[width=0.98\textwidth]{figures/multi_panel.pdf}
+\caption{Prediction for \tess\ amplitudes relative to the \kepler\ sample.}
+\label{fig:predictAmps}
+\end{figure}
 
@@
 
-\subsection{Measuring periods and amplitudes, which has been demonstrated}
-\subsection{There will be enough GKM stars}
-\subsection{The TESS precision will be adequate}
-Rapid rotators tend to exhibit larger amplitudes of lightcurve undulations than slower rotators.
-\subsection{Experience with the spectral technique on young stars}
-\subsection{Availability of archival spectroscopy}
+Measuring periods and amplitudes of \kepler\ and \tess\ lightcurves is a routine operation, with relatively low risk.
+The PI has experience determining amplitudes with a Fourier-based decomposition technique in \ktwo\ Campaign 2.
+The \tess\ mission was designed to target cool main sequence stars with an abundance of K and M dwarfs.
+This program will focus on the rapidly rotating stars ($P_{rot}<20$ days), which exhibit relatively large amplitudes ($A>2000$ ppm), illustrated in panel D of Figure \ref{fig:predictAmps}.
+This program will attempt---on a best effort basis---to measure the slowest rotating stars, which have periods much longer than the typical observation window of 27-54 days, and amplitudes comparable to the precision limits of \tess; these slow rotating sources must reside near the continuous viewing zone in the North Ecliptic Pole.
+
+One potential risk is the that the \kepler\ and \tess\ samples will possess statistical differences beyond their observation bands, confounding the...
 
@@
 
-The Pincipal Investigator (PI) will carry out the data analysis using custom scripts that build upon the heritage of the Kepler data analysis ecosystem, including \texttt{lightkurve}, \texttt{astropy}, \texttt{celerite}, \texttt{BombScargle}, \texttt{PyTorch}, and \texttt{bokeh}.
-The PI will first apply these scripts to Kepler/K2 lightcurves while in their current role as support scientist at the Kepler/K2 Guest Observer Office.
+The Principal Investigator (PI) will carry out the data analysis using custom scripts that build upon the heritage of the \kepler\ data analysis ecosystem, including \texttt{lightkurve}, \texttt{astropy}, \texttt{celerite}, \texttt{BombScargle}, \texttt{PyTorch}, and \texttt{bokeh}.
+The PI will first apply these scripts to Kepler/K2 lightcurves while in the role as support scientist at the Kepler/K2 Guest Observer Office.
 The PI will relocate to the University of Texas at Austin in February 2020 as a Research Fellow.
@@
 This tool will leverage the PI's familiarity with \texttt{bokeh} to aid understanding and rapid visual spotchecking of the multidimensional analysis output.
-Finally, a High Level Science Product (HLSP) will be delivered to MAST that contains the following information for each star found to possess detectable starspot-induced lightcurve modulation: 1) the period or periods, 2) an iteratively determined outlier mask of stellar flares, occultations, or spurious data artifacts, 3) a vector of Chebychev polynomial coefficients to account for any long-term normalization trends on timescales comparable to the observation window, and 4) an $N_{\mathrm{Fourier}}$ vector of Fourier harmonic amplitudes that reconstruct the periodic structures in the lightcurve to within the signal-to-noise ratio of the data.
+Finally, a High Level Science Product (HLSP) will be delivered to MAST that contains the following information for each star found to possess detectable starspot-induced lightcurve modulation: 1) the period or periods, 2) an iteratively determined outlier mask of stellar flares, occultations, or spurious data artifacts, 3) a background-star amplitude-dillution estimate, 4) a vector of Chebychev polynomial coefficients to account for any long-term normalization trends on timescales comparable to the observation window, and 5) a vector of Fourier harmonic amplitudes that reconstruct the periodic structures in the lightcurve to within the signal-to-noise ratio of the data.
 MGS will advise students in the writeup of the paper(s), authored and/or coauthored by students.

### 2019-03-14 proposals/TESS_GI_cycle02/tessgi_gully_cycle2.tex
Updated with content in all sections

@@
 
-\section{Scientific Justification}
-
-\subsection{Transit Light Source Effect}
+\section{Scientific Justification: The Transit Light Source Effect}
 The diminution of flux arising from occultations of the stellar disk contains information about the relative sizes of the occulting bodies.
@@
 
-\subsection{Kepler and TESS perceive different starspot contrast}
+\textbf{Kepler and TESS perceive different starspot contrasts.}s
 
@@
 
-\subsection{Spot-induced modulation amplitudes depend on contrast}
+\textbf{Spot-induced modulation amplitudes depend on contrast.}\\
 Rotationally-modulated lightcurves exhibit semi-sinusoidal undulations as starspots (or spot groups) enter and exit the projected stellar disk.
 The most-spotted projected hemisphere is directed towards the observer at the moment of the lightcurve minimum, and the least-spotted projected hemisphere at lightcurve maximum.
-The amplitude of the lightcurve therefore equals the difference in starspot coverage fractions $\Delta f_{spot}$ between the most-and-least spotted projected hemispheres \emph{weighted by the spot contrast}:
+The amplitude of the lightcurve therefore equals the difference in starspot coverage fractions $\Delta f_{spot} \equiv f_{max}-f_{min}$ between the most-and-least spotted projected hemispheres \emph{weighted by the spot contrast}:
 
@@
 
-\subsection{Measuring Kepler/K2 and TESS amplitudes: Statistical samples}
+Recent stellar surface inhomogeneity modeling \citep{2018ApJ...853..122R, 2018ApJ...868..143G} has yielded coarse scaling factors $\zeta$ for estimating the total starspot surface coverage fraction from the observed lightcurve amplitude $f_{spot}$:
+
+$$ p(f_{spot}) = \zeta \cdot A $$
+
+where $\zeta$ represents a probability distribution with values typically between 1 and 50, depending on the degree of assumed starspot longitudinal symmetry.
+Equipped with \tess-measured $A$ values, we can provide what practitioners seek: a per-star estimate of $f_{spot}$ and $T_{spot}$.
+
+\textbf{Measuring Kepler/K2 and TESS amplitudes with statistical samples}
+
 Ideally we would have simultaneouly observed the same targets with \kepler\ and \tess\ to measure the contrast for individual stars\footnote{Technically K2 Campaign 19 contemporaneously observed a subset of the same stars as \tess, but yielded few sources relevant to this discussion.}.
@@
 \begin{figure}
-\includegraphics[width=0.98\textwidth]{figures/multi_panel.pdf}
+\includegraphics[width=0.95\textwidth]{figures/multi_panel.pdf}
 \caption{Prediction for \tess\ amplitudes relative to the \kepler\ sample.}
@@
 Measuring periods and amplitudes of \kepler\ and \tess\ lightcurves is a routine operation, with relatively low risk.
-The PI has experience determining amplitudes with a Fourier-based decomposition technique in \ktwo\ Campaign 2.
+The PI has significant experience determining amplitudes with a Fourier-based decomposition technique in young stars observed in \ktwo.
+The derived periods and amplitudes need not be \emph{perfect}, so long as the biases are about the same in both the \kepler\ and \tess\ subsamples.
+
+The yield of spotted stars in \tess\ will number in the hundreds of thousands.
 The \tess\ mission was designed to target cool main sequence stars with an abundance of K and M dwarfs.
 This program will focus on the rapidly rotating stars ($P_{rot}<20$ days), which exhibit relatively large amplitudes ($A>2000$ ppm), illustrated in panel D of Figure \ref{fig:predictAmps}.
-This program will attempt---on a best effort basis---to measure the slowest rotating stars, which have periods much longer than the typical observation window of 27-54 days, and amplitudes comparable to the precision limits of \tess; these slow rotating sources must reside near the continuous viewing zone in the North Ecliptic Pole.
+This program will attempt---on a best effort basis---to measure the slowest rotating stars, which have periods much longer than the typical observation window of 27$-$54 days, and amplitudes comparable to the precision limits of \tess; these slow rotating sources must reside near the continuous viewing zone in the North Ecliptic Pole to have any hope of detecting spot modulation.
 
-One potential risk is the that the \kepler\ and \tess\ samples will possess statistical differences beyond their observation bands, confounding the...
+One potential risk to this program is the that the \kepler\ and \tess\ populations will be statistically different from each other in their unobserved properties, confounding the assumption that the subsamples' only difference is the observation bandpass.
+For example, metallicity, age, extinction, and stellar background contamination may all be different between the \kepler\ subsamples and \tess\ subsamples.
+Luckily, \tess\ Cycle 2 will have significant overlap with both the \kepler\ prime field (up to 3 sectors of overlap), and the northern edges of many \ktwo\ campaign fields.
+The total number of sources with significant overlap numbers in the tens-to-hundreds of thousands.
+This overlap sample offers an independent check on any statistical biases, since the exact same stars can be compared.
+The fact that observations are not contemporaneous add some \emph{cosmic variance}, but on average stars will enter and exit stellar activity cycles in equal measure, injecting per-star variance, but keeping the subsample mean amplitudes the same.
 
@@
 
-\subsection{Probabilistic lookup table for spot temperature and spot coverage}
-The probabilistic lookup table for spot temperature and filling factor will empower exoplanet researchers to constrain biases in exoplanet transit radii.
-
-\subsection{Direct estimate of spot coverage for selected high value targets}
-The spectroscopic analysis component of this work will add long-term value to the TESS legacy.
-
-\subsection{Trends in spot coverage with lightcurve periodic substructure}
-The adopted Fourier analysis technique naturally collates the relative amplitudes of periodic structures, which encode some additional information about the spatial distribution of stellar surface inhomogeneities, and/or differential rotation.
-The collated harmonics will enable exploratory analyses of trends with other observables, as recently proposed in the amount of time a spot-modulated \kepler\ lightcurve exhibits single or double dips \citep{2018ApJ...863..190B}.
+The main deliverable of this work will be a probabilistic assessment of the starspot temperature and starspot coverage faction for every star with a measured starspot modulation period and amplitude in \tess\.
+These probability distributions will empower exoplanet researchers to constrain biases in exoplanet transit radii.
+The probability distributions will reflect all the available information garnered from \tess, and reasonable assumptions about starspot geometrical distributions.
 

### 2019-03-14 proposals/TESS_GI_cycle02/tessgi_gully_cycle2.tex
pre-grammarly

@@
 Addtional correction factors account for limb darkening \citep{2002ApJ...580L.171M}.
-Starspots on a stellar surface can exhibit spatial segrations %\footnote{Sunspots reside on the Sun's mid-latitudes; starspots can exist near the stellar poles, as evinced in spatially resolved interferometry of nearby giant stars \citep{2016Natur.533..217R}.}
+Starspots on a stellar surface can exhibit spatial segregations
 that yield \emph{unocculted starspots}: spots that are present on the surface but are not traversed during transits.
@@
 
-\textbf{Kepler and TESS perceive different starspot contrasts.}s
+\textbf{Kepler and TESS perceive different starspot contrasts.}
 
 Starspots emit light, which is often parameterized as \emph{spot contrast}, $c$, where 0 means the spots are non-emitting, and $c=1$ means thes spot matches the specific intensity of the ambient photosphere.
-The sun
-
-The spot contrast is wavelength dependent ($c_\lambda$), and can be thought of as possessing a characteristic spot temperature, $T_{spot}<T_{phot}$ that parameterizes the wavelength dependence of the contrasts.
-
+Spot contrast is wavelength dependent ($c_\lambda$), and can be thought of as possessing a characteristic spot temperature, $T_{spot}<T_{phot}$ that parameterizes the wavelength dependence of the contrasts.
 The analysis technique proposed here leverages the different wavelength dependence of the \kepler\ and \tess\ system throughputs to directly measure the ratio of the spot contrasts in their respective bandpasses.
-
-$$\frac{c_{TESS}}{c_{Kepler}}(\mathrm{SpT},\;P_{\mathrm{rot}})$$
-
 The spot contrast ratio---evaluated at these two pairs of bandpasses---provides a unique solution\footnote{Operationally, integrating the specific intensity $I_\lambda(T_{spot})$ over the system throughputs produces the expected contrast ratios.
@@
 
-\textbf{Spot-induced modulation amplitudes depend on contrast.}\\
+\textbf{Spot-induced modulation amplitudes depend on contrast.}
 Rotationally-modulated lightcurves exhibit semi-sinusoidal undulations as starspots (or spot groups) enter and exit the projected stellar disk.
@@
 
-Recent stellar surface inhomogeneity modeling \citep{2018ApJ...853..122R, 2018ApJ...868..143G} has yielded coarse scaling factors $\zeta$ for estimating the total starspot surface coverage fraction from the observed lightcurve amplitude $f_{spot}$:
-
-$$ p(f_{spot}) = \zeta \cdot A $$
-
-where $\zeta$ represents a probability distribution with values typically between 1 and 50, depending on the degree of assumed starspot longitudinal symmetry.
+Recent stellar surface inhomogeneity modeling \citep{2018ApJ...853..122R, 2018ApJ...868..143G, 2018ApJ...865..142B} has yielded coarse scaling factors $\zeta$ for estimating the total starspot surface coverage fraction from the observed lightcurve amplitude $f_{spot}$, yielding $$ p(f_{spot}) = \zeta \cdot A $$, where $\zeta$ represents a probability distribution with values typically between 1 and 50, depending on the degree of assumed starspot longitudinal symmetry.
 Equipped with \tess-measured $A$ values, we can provide what practitioners seek: a per-star estimate of $f_{spot}$ and $T_{spot}$.
 
-\textbf{Measuring Kepler/K2 and TESS amplitudes with statistical samples}
+\textbf{Distinguishing Kepler/K2 and TESS amplitudes with statistical similar subsamples}
 
-Ideally we would have simultaneouly observed the same targets with \kepler\ and \tess\ to measure the contrast for individual stars\footnote{Technically K2 Campaign 19 contemporaneously observed a subset of the same stars as \tess, but yielded few sources relevant to this discussion.}.
-Since this observational strategy was not possible, we can construct statistical subsamples of stars that are identical in every way, except for their observation band.
+Ideally we would have simultaneouly observed the same targets with \kepler\ and \tess\ to measure the contrast for individual stars.
+Since this observational strategy was not possible for a large number of stars, we instead construct statistical subsamples of stars that are identical in every way, except for their observation band.
 
 The main two factors that affect spot-modulation lightcurve amplitudes are effective temperature (\emph{i.e.} mass), and rotation period (\emph{i.e.} stellar activity/magnetic fields) \citep{2014ApJS..211...24M}.
-Other important albeit unknwon factors like stellar inclination angle and multiplicity are statistically identical between the \tess\ and \kepler\ samples.
+Other important-albeit-unknown factors like stellar inclination angle and multiplicity are statistically identical between the \tess\ and \kepler\ samples.
 We therefore propose to follow \citet{2014ApJS..211...24M} by constructing plots of amplitude versus period, subdivided into bins of effective temperature.
@@
 
-\subsection{Stellar surface structures as a gateway for brown dwarf cloud modeling}
+Finally, stellar surface structures serve as a stepping stone for brown dwarf cloud modeling, a poorly understood phenomenon that will inform our understand of exoplanet spectroscopy.
+The proposed work will demonstrate how to gain insights about probable surface structures in the presence of significant geoemtrical degeneracies.
 

### 2019-03-14 proposals/TESS_GI_cycle02/tessgi_gully_cycle2.tex
post grammarly

@@
 A commonsense exoplanet heuristic dictates that the relative transit depth $D$ equals the ratio of the bodies' projected areas ($R_p^2/R_\star^2$).
-Addtional correction factors account for limb darkening \citep{2002ApJ...580L.171M}.
+Additional correction factors account for limb darkening \citep{2002ApJ...580L.171M}.
 Starspots on a stellar surface can exhibit spatial segregations
@@
 New theoretical models \citep{2018ApJ...853..122R} and observational evidence \citep{2016MNRAS.463.2494F} indicate that starspot covering fractions are large enough to bias transit depths beyond the level of precision of \kepler\ and \tess.
-This so-called ``Transit Light Source Effect'' (TLSE) injects significant bias into derived exoplanet \emph{densities}, since the radius error propagates into volume error with a factor of 3.
+This so-called ``Transit Light Source Effect'' (TLSE) injects significant bias into derived exoplanet \emph{densities} since the radius error propagates into volume error with a factor of 3.
 Therefore, TLSE threatens placement and interpretation of the exoplanet mass-radius relation science requirements of the \tess\ mission.
@@
 
-Starspots emit light, which is often parameterized as \emph{spot contrast}, $c$, where 0 means the spots are non-emitting, and $c=1$ means thes spot matches the specific intensity of the ambient photosphere.
-Spot contrast is wavelength dependent ($c_\lambda$), and can be thought of as possessing a characteristic spot temperature, $T_{spot}<T_{phot}$ that parameterizes the wavelength dependence of the contrasts.
+Starspots emit light, which is often parameterized as \emph{spot contrast}, $c$, where 0 means the spots are non-emitting, and $c=1$ means the spot matches the specific intensity of the ambient photosphere.
+Spot contrast is wavelength dependent ($c_\lambda$) and can be thought of as possessing a characteristic spot temperature, $T_{spot}<T_{phot}$ that parameterizes the wavelength dependence of the contrasts.
 The analysis technique proposed here leverages the different wavelength dependence of the \kepler\ and \tess\ system throughputs to directly measure the ratio of the spot contrasts in their respective bandpasses.
@@
 The most-spotted projected hemisphere is directed towards the observer at the moment of the lightcurve minimum, and the least-spotted projected hemisphere at lightcurve maximum.
-The amplitude of the lightcurve therefore equals the difference in starspot coverage fractions $\Delta f_{spot} \equiv f_{max}-f_{min}$ between the most-and-least spotted projected hemispheres \emph{weighted by the spot contrast}:
+The amplitude of the lightcurve, therefore, equals the difference in starspot coverage fractions $\Delta f_{spot} \equiv f_{max}-f_{min}$ between the most-and-least spotted projected hemispheres \emph{weighted by the spot contrast}:
 
@@
 
-Ideally we would have simultaneouly observed the same targets with \kepler\ and \tess\ to measure the contrast for individual stars.
+Ideally, we would have simultaneously observed the same targets with \kepler\ and \tess\ to measure the contrast for individual stars.
 Since this observational strategy was not possible for a large number of stars, we instead construct statistical subsamples of stars that are identical in every way, except for their observation band.
@@
 Other important-albeit-unknown factors like stellar inclination angle and multiplicity are statistically identical between the \tess\ and \kepler\ samples.
-We therefore propose to follow \citet{2014ApJS..211...24M} by constructing plots of amplitude versus period, subdivided into bins of effective temperature.
+We, therefore, propose to follow \citet{2014ApJS..211...24M} by constructing plots of amplitude versus period, subdivided into bins of effective temperature.
 The two samples should look identical in these plots, but the locus of \tess\ sources are expected to fall below the \kepler\ sample, as in Figure \ref{fig:predictAmps}.
@@
 For example, metallicity, age, extinction, and stellar background contamination may all be different between the \kepler\ subsamples and \tess\ subsamples.
-Luckily, \tess\ Cycle 2 will have significant overlap with both the \kepler\ prime field (up to 3 sectors of overlap), and the northern edges of many \ktwo\ campaign fields.
+Luckily, \tess\ Cycle 2 will have significant overlap with both the \kepler\ prime field (up to 3 sectors of overlap) and the northern edges of many \ktwo\ campaign fields.
 The total number of sources with significant overlap numbers in the tens-to-hundreds of thousands.
-This overlap sample offers an independent check on any statistical biases, since the exact same stars can be compared.
+This overlap sample offers an independent check on any statistical biases since the exact same stars can be compared.
 The fact that observations are not contemporaneous add some \emph{cosmic variance}, but on average stars will enter and exit stellar activity cycles in equal measure, injecting per-star variance, but keeping the subsample mean amplitudes the same.
@@
 
-Finally, stellar surface structures serve as a stepping stone for brown dwarf cloud modeling, a poorly understood phenomenon that will inform our understand of exoplanet spectroscopy.
-The proposed work will demonstrate how to gain insights about probable surface structures in the presence of significant geoemtrical degeneracies.
+Finally, stellar surface structures serve as a stepping stone for brown dwarf cloud modeling, a poorly understood phenomenon that will inform our understanding of exoplanet spectroscopy.
+The proposed work will demonstrate how to gain insights about probable surface structures in the presence of significant geometrical degeneracies.
 
@@
 The Principal Investigator (PI) will carry out the data analysis using custom scripts that build upon the heritage of the \kepler\ data analysis ecosystem, including \texttt{lightkurve}, \texttt{astropy}, \texttt{celerite}, \texttt{BombScargle}, \texttt{PyTorch}, and \texttt{bokeh}.
-The PI will first apply these scripts to Kepler/K2 lightcurves while in the role as support scientist at the Kepler/K2 Guest Observer Office.
+The PI will first apply these scripts to Kepler/K2 lightcurves while in the role as a support scientist at the Kepler/K2 Guest Observer Office.
 The PI will relocate to the University of Texas at Austin in February 2020 as a Research Fellow.
@@
 MGS will provide students with an interactive \texttt{bokeh} tool to visualize the Fourier decomposition of a periodic lightcurve into its constituent components.
-This tool will leverage the PI's familiarity with \texttt{bokeh} to aid understanding and rapid visual spotchecking of the multidimensional analysis output.
-Finally, a High Level Science Product (HLSP) will be delivered to MAST that contains the following information for each star found to possess detectable starspot-induced lightcurve modulation: 1) the period or periods, 2) an iteratively determined outlier mask of stellar flares, occultations, or spurious data artifacts, 3) a background-star amplitude-dillution estimate, 4) a vector of Chebychev polynomial coefficients to account for any long-term normalization trends on timescales comparable to the observation window, and 5) a vector of Fourier harmonic amplitudes that reconstruct the periodic structures in the lightcurve to within the signal-to-noise ratio of the data.
+This tool will leverage the PI's familiarity with \texttt{bokeh} to aid understanding and rapid visual spot-checking of the multidimensional analysis output.
+Finally, a High-Level Science Product (HLSP) will be delivered to MAST that contains the following information for each star found to possess detectable starspot-induced lightcurve modulation: 1) the period or periods, 2) an iteratively determined outlier mask of stellar flares, occultations, or spurious data artifacts, 3) a background-star amplitude-dilution estimate, 4) a vector of Chebychev polynomial coefficients to account for any long-term normalization trends on timescales comparable to the observation window, and 5) a vector of Fourier harmonic amplitudes that reconstruct the periodic structures in the lightcurve to within the signal-to-noise ratio of the data.
 MGS will advise students in the writeup of the paper(s), authored and/or coauthored by students.

### 2019-03-14 proposals/TESS_GI_cycle02/tessgi_gully_cycle2.tex
pre CVM feedback

@@
 \bibliographystyle{plainnat}
+%\bibliographystyle{yahapj}
 %\usepackage{cleveref}
@@
 
+\usepackage{paralist}
+\makeatletter
+\newcommand{\mybullet}{%
+    \ifnum\value{enumi}=1
+    \else
+        \textbullet
+    \fi
+}
+\makeatother
+\renewenvironment{thebibliography}[1]{\let\par\relax%
+  \section*{\refname}\inparaenum[\mybullet]}{\endinparaenum}
+\let\oldbibitem\bibitem
+\renewcommand{\bibitem}{\item \oldbibitem}
+\include{JournalAbbr}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %% ENTER TITLE OF PROPOSAL BELOW THIS LINE
-Starspots confound derived stellar properties and a strategy for probabilistic mitigation
+Mitigating exoplanet density biases from wavelength dependence of spot contrast
 %%
@@
 
+Recent work has shown that ignoring starspots biases measurements of exoplanet radii \citep{2018ApJ...853..122R}.
 Here we propose a strategy for mitigating starspot-induced exoplanet radius biases, based on a large sample of spotted stars observed in \tess\ Cycle 2.
@@
 
-\begin{figure}
+\begin{figure}[hbt!]
 \includegraphics[width=0.95\textwidth]{figures/multi_panel.pdf}
-\caption{Prediction for \tess\ amplitudes relative to the \kepler\ sample.}
+\caption{\texttt{A)} Schematic of a planet ($R_p = 0.1 R_\star$) transiting a star with unocculted starspots as the star moves from left to right.
+\texttt{B)} The resulting transit depth shows a systematic bias and a wavelength dependence due to starspots.
+\texttt{C)} Illustration of \texttt{D)} Prediction for \tess\ amplitudes (colorful contours) relative to the \kepler\ sample (gray shaded background contours and black points) for $4000 < T_{phot} (K) < 4500$ \citep{2014ApJS..211...24M}.}
 \label{fig:predictAmps}

### 2021-01-19 proposals/TESS_GI_cycle04/tessgi_gully_cycle2.tex
Start of TESS GI Cycle 4

--- 2019-03-14 proposals/TESS_GI_cycle02/tessgi_gully_cycle2.tex
@@
 %% ENTER TITLE OF PROPOSAL BELOW THIS LINE
-Mitigating exoplanet density biases from wavelength dependence of spot contrast
+Mitigating exoplanet radius biases from wavelength dependence of starspot contrast
 %%
@@
 
-Recent work has shown that ignoring starspots biases measurements of exoplanet radii \citep{2018ApJ...853..122R}.
+Recent work has shown that ignoring starspots systematically biases measurements of exoplanet radii \citep{2018ApJ...853..122R}.
 Here we propose a strategy for mitigating starspot-induced exoplanet radius biases, based on a large sample of spotted stars observed in \tess\ Cycle 2.
@@
 No such reliable lookup table presently exists for main sequence stars, owing to the inability of differential lightcurves to account for \emph{total} starspot coverage fractions \citep{2018ApJ...865..142B}.
-
 This program will leverage the \tess\ Cycle 2 overlap with the \kepler\ prime field and portions of \ktwo\ fields.
@@
 \section{Scientific Justification: The Transit Light Source Effect}
-The diminution of flux arising from occultations of the stellar disk contains information about the relative sizes of the occulting bodies.
+The diminution of flux arising from occultations of the stellar disk contains information the relative sizes of the occulting bodies.
 A commonsense exoplanet heuristic dictates that the relative transit depth $D$ equals the ratio of the bodies' projected areas ($R_p^2/R_\star^2$).
@@
 New theoretical models \citep{2018ApJ...853..122R} and observational evidence \citep{2016MNRAS.463.2494F} indicate that starspot covering fractions are large enough to bias transit depths beyond the level of precision of \kepler\ and \tess.
-This so-called ``Transit Light Source Effect'' (TLSE) injects significant bias into derived exoplanet \emph{densities} since the radius error propagates into volume error with a factor of 3.
+This so-called ``Transit Light Source Effect'' (TLSE) can bias derived exoplanet radii up to 17\% in the \tess\ bandpass, and can bias derived exoplanet \emph{densities} up to 50\%, since the radius error propagates into volume error with a factor of 3.
 Therefore, TLSE threatens placement and interpretation of the exoplanet mass-radius relation science requirements of the \tess\ mission.
-The interpretation of \jwst\ exoplanet transmission spectroscopy is also at risk without improved knowledge of the level of the stellar photospheric contamination \citep{2019AJ....157...11W}.
+The interpretation of \jwst\ exoplanet transmission spectroscopy is also at risk \citep{2019AJ....157...11W}.
 
@@
 The most-spotted projected hemisphere is directed towards the observer at the moment of the lightcurve minimum, and the least-spotted projected hemisphere at lightcurve maximum.
-The amplitude of the lightcurve, therefore, equals the difference in starspot coverage fractions $\Delta f_{spot} \equiv f_{max}-f_{min}$ between the most-and-least spotted projected hemispheres \emph{weighted by the spot contrast}:
-
-$$ A_\lambda = \Delta f_{spot} \cdot (1-c_\lambda)$$
-
+The amplitude of the lightcurve, therefore, equals the difference in starspot coverage fractions $\Delta f_{spot} \equiv f_{max}-f_{min}$ between the most-and-least spotted projected hemispheres weighted by the spot contrast: $ A_\lambda = \Delta f_{spot} \cdot (1-c_\lambda)$.
 Therefore the ratio of \tess\ to \kepler\ amplitudes yields a constraint independent of the actual spot coverage fraction:
@@
 
-Recent stellar surface inhomogeneity modeling \citep{2018ApJ...853..122R, 2018ApJ...868..143G, 2018ApJ...865..142B} has yielded coarse scaling factors $\zeta$ for estimating the total starspot surface coverage fraction from the observed lightcurve amplitude $f_{spot}$, yielding $$ p(f_{spot}) = \zeta \cdot A $$, where $\zeta$ represents a probability distribution with values typically between 1 and 50, depending on the degree of assumed starspot longitudinal symmetry.
-Equipped with \tess-measured $A$ values, we can provide what practitioners seek: a per-star estimate of $f_{spot}$ and $T_{spot}$.
+Recent stellar surface inhomogeneity modeling \citep{2018ApJ...853..122R, 2018ApJ...868..143G, 2018ApJ...865..142B} has yielded coarse scaling factors $\zeta$ for estimating the total starspot surface coverage fraction from the observed lightcurve amplitude $f_{spot}$, yielding $ p(f_{spot}) = \zeta \cdot A $, where $\zeta$ represents a probability distribution with values typically between 1 and 50, depending on the degree of assumed starspot longitudinal symmetry.
+Equipped solely with \tess-measured lightcurve amplitudes, we can provide what practitioners seek: a per-star estimate of $f_{spot}$ and $T_{spot}$.
 
-\textbf{Distinguishing Kepler/K2 and TESS amplitudes with statistical similar subsamples}
+\textbf{Distinguishing Kepler/K2 and TESS amplitudes with statistically similar samples}
 
@@
 We, therefore, propose to follow \citet{2014ApJS..211...24M} by constructing plots of amplitude versus period, subdivided into bins of effective temperature.
-The two samples should look identical in these plots, but the locus of \tess\ sources are expected to fall below the \kepler\ sample, as in Figure \ref{fig:predictAmps}.
+The two samples should look identical in these plots, but the locus of \tess\ sources is expected to fall below the \kepler\ sample, as in panel D of Figure \ref{fig:predictAmps}.
 
@@
 \texttt{B)} The resulting transit depth shows a systematic bias and a wavelength dependence due to starspots.
-\texttt{C)} Illustration of \texttt{D)} Prediction for \tess\ amplitudes (colorful contours) relative to the \kepler\ sample (gray shaded background contours and black points) for $4000 < T_{phot} (K) < 4500$ \citep{2014ApJS..211...24M}.}
+\texttt{C)} Illustration of TLSE, \texttt{D)} Prediction for \tess\ amplitudes (colorful contours) relative to the \kepler\ sample (gray shaded background contours and black points) for $4000 < T_{phot} (K) < 4500$ \citep{2014ApJS..211...24M}.}
 \label{fig:predictAmps}
 \end{figure}
+
+\tess\ Cycle 2 will have significant overlap with both the \kepler\ prime field (up to 3 sectors of overlap) and 16438 source in the northern edges of \ktwo\ campaign fields (K.
+Col\'on, priv.
+comm.).
+This overlap sample offers an independent check on any statistical biases since the exact same stars can be compared.
+The fact that observations are not contemporaneous adds some \emph{cosmic variance}, but on average stars will enter and exit stellar activity cycles in equal measure, injecting per-star variance, but keeping the subsample mean amplitudes the same.
 
@@
 
-One potential risk to this program is the that the \kepler\ and \tess\ populations will be statistically different from each other in their unobserved properties, confounding the assumption that the subsamples' only difference is the observation bandpass.
-For example, metallicity, age, extinction, and stellar background contamination may all be different between the \kepler\ subsamples and \tess\ subsamples.
-Luckily, \tess\ Cycle 2 will have significant overlap with both the \kepler\ prime field (up to 3 sectors of overlap) and the northern edges of many \ktwo\ campaign fields.
-The total number of sources with significant overlap numbers in the tens-to-hundreds of thousands.
-This overlap sample offers an independent check on any statistical biases since the exact same stars can be compared.
-The fact that observations are not contemporaneous add some \emph{cosmic variance}, but on average stars will enter and exit stellar activity cycles in equal measure, injecting per-star variance, but keeping the subsample mean amplitudes the same.
+One potential risk to this program is the that the \kepler\ and \tess\ populations will be statistically different from each other in their unobserved properties, confounding the assumption that the samples' only difference is the observation bandpass.
+For example, metallicity, age, extinction, and stellar background contamination may all be different between the \kepler\ samples and \tess\ samples.
+This risk is alleviated by the subsample of \kepler\ sources reobserved in \tess\ Cycle 2.
 
@@
 
-The main deliverable of this work will be a probabilistic assessment of the starspot temperature and starspot coverage faction for every star with a measured starspot modulation period and amplitude in \tess\.
-These probability distributions will empower exoplanet researchers to constrain biases in exoplanet transit radii.
+The main deliverable of this work will be a probabilistic assessment of the starspot temperature and starspot coverage faction for every star with a measured starspot modulation period and amplitude in \tess.
+These probability distributions will empower exoplanet researchers to correct for biases in exoplanet transit radii.
 The probability distributions will reflect all the available information garnered from \tess, and reasonable assumptions about starspot geometrical distributions.
 
-Finally, stellar surface structures serve as a stepping stone for brown dwarf cloud modeling, a poorly understood phenomenon that will inform our understanding of exoplanet spectroscopy.
+Stellar surface structures serve as a stepping stone for brown dwarf cloud modeling, a poorly understood phenomenon that will inform our understanding of exoplanet spectroscopy.
 The proposed work will demonstrate how to gain insights about probable surface structures in the presence of significant geometrical degeneracies.
+
+The starspot properties are also of intrinsic astrophysical interest.
+Trends in $T_{spot}$ with rotation rate and mass can reveal important clues to stellar activity, magnetic fields, and stellar atmospheres.
+This work will directly address starspot-induced scatter and bias in isochronal-derived stellar ages \citep{2015ApJ...807..174S, 2017ApJ...836..200G}.
 
@@
 
-The Principal Investigator (PI) will carry out the data analysis using custom scripts that build upon the heritage of the \kepler\ data analysis ecosystem, including \texttt{lightkurve}, \texttt{astropy}, \texttt{celerite}, \texttt{BombScargle}, \texttt{PyTorch}, and \texttt{bokeh}.
+PI Gully-Santiago will carry out the data analysis using custom scripts that build upon the heritage of the \kepler\ data analysis ecosystem, including \texttt{lightkurve}, \texttt{astropy}, \texttt{celerite}, \texttt{BombScargle}, \texttt{PyTorch}, and \texttt{bokeh}.
 The PI will first apply these scripts to Kepler/K2 lightcurves while in the role as a support scientist at the Kepler/K2 Guest Observer Office.
 The PI will relocate to the University of Texas at Austin in February 2020 as a Research Fellow.
-UT Austin undergraduate and graduate students will lead the application of TESS-adapted scripts to \tess\ FFI lightcurves.
-\tess-adapted scripts will leverage \texttt{TESS-cut} and \texttt{elenor} to produce lightcurves from FFIs.
-Half of the funding will go to summer 2020 student salaries, and half will go to support the PI.
+UT Austin undergraduate and graduate students will lead the application of TESS-adapted scripts to \tess\ FFI lightcurves with \texttt{TESS-cut} and \texttt{elenor}.
+Half of the funding will go to 1 undergraduate and 1 graduate student Summer 2020 salary, and 0.3 FTE to support the PI.
 All reproducible scripts---including Jupyter Notebook tutorials---will be publicly available and developed openly on GitHub.
-MGS will provide students with an interactive \texttt{bokeh} tool to visualize the Fourier decomposition of a periodic lightcurve into its constituent components.
-This tool will leverage the PI's familiarity with \texttt{bokeh} to aid understanding and rapid visual spot-checking of the multidimensional analysis output.
-Finally, a High-Level Science Product (HLSP) will be delivered to MAST that contains the following information for each star found to possess detectable starspot-induced lightcurve modulation: 1) the period or periods, 2) an iteratively determined outlier mask of stellar flares, occultations, or spurious data artifacts, 3) a background-star amplitude-dilution estimate, 4) a vector of Chebychev polynomial coefficients to account for any long-term normalization trends on timescales comparable to the observation window, and 5) a vector of Fourier harmonic amplitudes that reconstruct the periodic structures in the lightcurve to within the signal-to-noise ratio of the data.
-MGS will advise students in the writeup of the paper(s), authored and/or coauthored by students.
+MGS will provide students with an interactive \texttt{bokeh} tool to visualize periodic lightcurve structures.
+This tool will leverage the PI's familiarity with \texttt{bokeh} to aid rapid visual spot-checking of the multidimensional analysis output.
+Finally, a High-Level Science Product (HLSP) will be delivered to MAST that contains the following information for each star found to possess detectable starspot-induced lightcurve modulation: 1) the period or periods, 2) an iteratively determined outlier mask of stellar flares, occultations, or spurious data artifacts, 3) a background-star amplitude-dilution estimate, 4) a probability distribution for the spot temperature and coverage fraction, 5) addition model fitting metadata.
+MGS will co-advise---with Co-I C.
+Morley---students in the writeup of the paper(s), authored and/or coauthored by students.
 
+{\footnotesize
 \bibliography{tessgi_gully_cycle2}
+}
 

### 2021-01-19 proposals/TESS_GI_cycle04/tessgi_gully_cycle4.tex
Refresh for 2021 Cycle 4 proposal

--- 2021-01-19 proposals/TESS_GI_cycle04/tessgi_gully_cycle2.tex
@@
 \usepackage{aas_macros}
-\usepackage[comma,authoryear]{natbib}
-\bibpunct{(}{)}{;}{a}{}{;}
+%\usepackage[comma,authoryear]{natbib}
+%\bibpunct{(}{)}{;}{a}{}{;}
 
@@
 
+\usepackage{fontspec}
+\usepackage[T1]{fontenc}
+\usepackage{newtxsf}
+\setmainfont{Fira Sans Book}[Scale=0.95]
+
 \usepackage[backref,breaklinks,colorlinks,urlcolor=blue,citecolor=blue,linkcolor=blue]{hyperref}
-\bibliographystyle{plainnat}
+\bibliographystyle{plain}
 %\bibliographystyle{yahapj}
-%\usepackage{cleveref}
+\usepackage{cleveref}
 \input{mgs_setup}
@@
 
-\usepackage{paralist}
-\makeatletter
-\newcommand{\mybullet}{%
-    \ifnum\value{enumi}=1
-    \else
-        \textbullet
-    \fi
-}
-\makeatother
-\renewenvironment{thebibliography}[1]{\let\par\relax%
-  \section*{\refname}\inparaenum[\mybullet]}{\endinparaenum}
-\let\oldbibitem\bibitem
-\renewcommand{\bibitem}{\item \oldbibitem}
-\include{JournalAbbr}
+%\include{JournalAbbr}
 
@@
 \pagenumbering{arabic}
-\setlength{\bibsep}{0.0pt}
-\let\oldbibliography\thebibliography
-\renewcommand{\thebibliography}[1]{%
-  \oldbibliography{#1}%
-  \setlength{\itemsep}{0pt}%
-}
 
@@
 
-Recent work has shown that ignoring starspots systematically biases measurements of exoplanet radii \citep{2018ApJ...853..122R}.
+Recent work has shown that ignoring starspots systematically biases measurements of exoplanet radii \cite{2018ApJ...853..122R}.
 Here we propose a strategy for mitigating starspot-induced exoplanet radius biases, based on a large sample of spotted stars observed in \tess\ Cycle 2.
@@
 This mapping can then serve as a ``lookup table'' for practitioners to gauge the extent to which a given source is likely to suffer from starspot-induced exoplanet radius biases.
-No such reliable lookup table presently exists for main sequence stars, owing to the inability of differential lightcurves to account for \emph{total} starspot coverage fractions \citep{2018ApJ...865..142B}.
+No such reliable lookup table presently exists for main sequence stars, owing to the inability of differential lightcurves to account for \emph{total} starspot coverage fractions \cite{2018ApJ...865..142B}.
 This program will leverage the \tess\ Cycle 2 overlap with the \kepler\ prime field and portions of \ktwo\ fields.
@@
 A commonsense exoplanet heuristic dictates that the relative transit depth $D$ equals the ratio of the bodies' projected areas ($R_p^2/R_\star^2$).
-Additional correction factors account for limb darkening \citep{2002ApJ...580L.171M}.
+Additional correction factors account for limb darkening \cite{2002ApJ...580L.171M}.
 Starspots on a stellar surface can exhibit spatial segregations
 that yield \emph{unocculted starspots}: spots that are present on the surface but are not traversed during transits.
-These unocculted starspots confound the mapping of transit depth to planet radius: a planet ``looks bigger'' than it really is if it blocks the brighter-on-average flux of the stellar disk \citep{2018AJ....156...91M}.
+These unocculted starspots confound the mapping of transit depth to planet radius: a planet ``looks bigger'' than it really is if it blocks the brighter-on-average flux of the stellar disk \cite{2018AJ....156...91M}.
 \textbf{You cannot measure a planet radius without making some assumptions about the extent, contrast, and distribution of starspots.}
@@
 The prevailing assumption within the exoplanet community has been that starspot covering fractions are small enough---as in Sunspots---to be ignored.
-New theoretical models \citep{2018ApJ...853..122R} and observational evidence \citep{2016MNRAS.463.2494F} indicate that starspot covering fractions are large enough to bias transit depths beyond the level of precision of \kepler\ and \tess.
+New theoretical models \cite{2018ApJ...853..122R} and observational evidence \cite{2016MNRAS.463.2494F} indicate that starspot covering fractions are large enough to bias transit depths beyond the level of precision of \kepler\ and \tess.
 This so-called ``Transit Light Source Effect'' (TLSE) can bias derived exoplanet radii up to 17\% in the \tess\ bandpass, and can bias derived exoplanet \emph{densities} up to 50\%, since the radius error propagates into volume error with a factor of 3.
 Therefore, TLSE threatens placement and interpretation of the exoplanet mass-radius relation science requirements of the \tess\ mission.
-The interpretation of \jwst\ exoplanet transmission spectroscopy is also at risk \citep{2019AJ....157...11W}.
+The interpretation of \jwst\ exoplanet transmission spectroscopy is also at risk \cite{2019AJ....157...11W}.
 
@@
 The spot contrast ratio---evaluated at these two pairs of bandpasses---provides a unique solution\footnote{Operationally, integrating the specific intensity $I_\lambda(T_{spot})$ over the system throughputs produces the expected contrast ratios.
-For now we assume $I_\lambda(T_{spot})$ can be approximated as a cooler stellar photosphere, such as the \texttt{PHOENIX} model grid \citep{2013A&A...553A...6H}, an inexact but adequate assumption for the large bandpasses.} to $T_{spot}(\mathrm{SpT},\;P_{\mathrm{rot}})$.
+For now we assume $I_\lambda(T_{spot})$ can be approximated as a cooler stellar photosphere, such as the \texttt{PHOENIX} model grid \cite{2013A&A...553A...6H}, an inexact but adequate assumption for the large bandpasses.} to $T_{spot}(\mathrm{SpT},\;P_{\mathrm{rot}})$.
 
@@
 
-Recent stellar surface inhomogeneity modeling \citep{2018ApJ...853..122R, 2018ApJ...868..143G, 2018ApJ...865..142B} has yielded coarse scaling factors $\zeta$ for estimating the total starspot surface coverage fraction from the observed lightcurve amplitude $f_{spot}$, yielding $ p(f_{spot}) = \zeta \cdot A $, where $\zeta$ represents a probability distribution with values typically between 1 and 50, depending on the degree of assumed starspot longitudinal symmetry.
+Recent stellar surface inhomogeneity modeling \cite{2018ApJ...853..122R, 2018ApJ...868..143G, 2018ApJ...865..142B} has yielded coarse scaling factors $\zeta$ for estimating the total starspot surface coverage fraction from the observed lightcurve amplitude $f_{spot}$, yielding $ p(f_{spot}) = \zeta \cdot A $, where $\zeta$ represents a probability distribution with values typically between 1 and 50, depending on the degree of assumed starspot longitudinal symmetry.
 Equipped solely with \tess-measured lightcurve amplitudes, we can provide what practitioners seek: a per-star estimate of $f_{spot}$ and $T_{spot}$.
@@
 
-The main two factors that affect spot-modulation lightcurve amplitudes are effective temperature (\emph{i.e.} mass), and rotation period (\emph{i.e.} stellar activity/magnetic fields) \citep{2014ApJS..211...24M}.
+The main two factors that affect spot-modulation lightcurve amplitudes are effective temperature (\emph{i.e.} mass), and rotation period (\emph{i.e.} stellar activity/magnetic fields) \cite{2014ApJS..211...24M}.
 Other important-albeit-unknown factors like stellar inclination angle and multiplicity are statistically identical between the \tess\ and \kepler\ samples.
-We, therefore, propose to follow \citet{2014ApJS..211...24M} by constructing plots of amplitude versus period, subdivided into bins of effective temperature.
+We, therefore, propose to follow \cite{2014ApJS..211...24M} by constructing plots of amplitude versus period, subdivided into bins of effective temperature.
 The two samples should look identical in these plots, but the locus of \tess\ sources is expected to fall below the \kepler\ sample, as in panel D of Figure \ref{fig:predictAmps}.
@@
 \texttt{B)} The resulting transit depth shows a systematic bias and a wavelength dependence due to starspots.
-\texttt{C)} Illustration of TLSE, \texttt{D)} Prediction for \tess\ amplitudes (colorful contours) relative to the \kepler\ sample (gray shaded background contours and black points) for $4000 < T_{phot} (K) < 4500$ \citep{2014ApJS..211...24M}.}
+\texttt{C)} Illustration of TLSE, \texttt{D)} Prediction for \tess\ amplitudes (colorful contours) relative to the \kepler\ sample (gray shaded background contours and black points) for $4000 < T_{phot} (K) < 4500$ \cite{2014ApJS..211...24M}.}
 \label{fig:predictAmps}
@@
 Trends in $T_{spot}$ with rotation rate and mass can reveal important clues to stellar activity, magnetic fields, and stellar atmospheres.
-This work will directly address starspot-induced scatter and bias in isochronal-derived stellar ages \citep{2015ApJ...807..174S, 2017ApJ...836..200G}.
+This work will directly address starspot-induced scatter and bias in isochronal-derived stellar ages \cite{2015ApJ...807..174S, 2017ApJ...836..200G}.
 
@@
 
-{\footnotesize
+\newpage
+
 \bibliography{tessgi_gully_cycle2}
-}
 

### 2021-01-19 proposals/TESS_GI_cycle04/tessgi_gully_cycle4.tex
Bare minimum changes to make it compliant with Cycle 4

@@
-%% LaTeX template for the science justification & technical
-%% feasibility to be submitted as part of a TESS Guest Investigator
-%% Program proposal. This template is based on the proposal template
-%% used by the NuSTAR mission.
-%%
-%% TESS Guest Investigator Proposal Cycle 2 template
-%% V1.0
-%% 2017-08-04
-%% V1.1
-%% 2019-02-07
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%% DOCUMENT FORMAT %%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%% The default font was chosen to be easily readable while allowing
-%% sufficient material to be included.
-
-%% Please note that the proposal will be printed on US Letter size paper,
-%% 8.5 in x 11 in, and that formatting the text for other sizes will
-%% generally cause layout problems and may result in text being cut
-%% off near the edges. PLEASE DO NOT CHANGE THE 'LETTERPAPER' OPTION
-%% IN THE DOCUMENTCLASS COMMAND.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%% Default format: 11pt single column %%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%% NOTE: NASA ROSES requires body font size to be no smaller than 15
-%% characters per inch (equivalent to Times Roman 12 point).
-%%
-%% Minimum margin size is 1 inch from top, bottom, and sides.
-
-\documentclass[letterpaper,11pt]{article}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%% HOW TO INCLUDE FIGURES %%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%% Please see the ``Included packages'' section below.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%% Included packages %%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\documentclass[letterpaper,12pt]{article}
 
@@
 \usepackage{aas_macros}
-%\usepackage[comma,authoryear]{natbib}
-%\bibpunct{(}{)}{;}{a}{}{;}
 
@@
 \bibliographystyle{plain}
-%\bibliographystyle{yahapj}
 \usepackage{cleveref}
-\input{mgs_setup}
-
-%% Feel free to modify the included packages list to use your
-%% favorite packages.
-
-%% In the graphics and graphicx packages, Postscript and eps figures
-%% can be included using the \includegraphics command. The graphics
-%% package is part of standard LaTeX2e and provides a basic way of including a
-%% figure. The graphicx package is not standard, but extends the
-%% \includegraphics command to make it more user-friendly. If graphicx
-%% is not available on your system please remove it from the list of
-%% included packages above.
-
-%% Syntax:
-%% In the graphics package:
-%%
-%% \begin{figure}
-%% \includegraphics[llx,lly][urx,ury]{file}
-%% \end{figure}
-%%
-%% where ll denotes 'lower left' and ur 'upper right' and the x and y
-%% values are the coordinates of the PostScript bounding box in
-%% points. There are 72 points in an inch.
-%%
-%% In the graphicx package:
-%%
-%% \begin{figure}
-%% \includegraphics[key=val,key=val,...]{file}
-%% \end{figure}
-%%
-%% where some of the useful keys are: angle, width, height,
-%% keepaspectratio (='true' or 'false') and scale. Bounding box values
-%% can be given as [bb=llx lly urx ury].
-%%
-%% In either case you have to use LaTeX figure placement commands to
-%% position the figure on the page; \includegraphics will not do
-%% that. Both these commands also have other options that are listed
-%% in the LaTeX manual (for the graphics package) and in 'The LaTeX
-%% Graphics Companion' (for the graphicx package).
 
@@
 
-%\include{JournalAbbr}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-Recent work has shown that ignoring starspots systematically biases measurements of exoplanet radii \cite{2018ApJ...853..122R}.
-Here we propose a strategy for mitigating starspot-induced exoplanet radius biases, based on a large sample of spotted stars observed in \tess\ Cycle 2.
+Starspots bias percieved exoplanet radii \cite{2018ApJ...853..122R}.
+Here we propose a systematic approach to quantify and mitigate starspot-induced exoplanet radius biases based on a large sample of spotted stars to be observed in \tess\ Cycle 4.
 The approach will deliver constraints on both the typical starspot coverage fraction and the typical starspot temperature contrast, as a function of spectral type and rotation rate.
@@
 No such reliable lookup table presently exists for main sequence stars, owing to the inability of differential lightcurves to account for \emph{total} starspot coverage fractions \cite{2018ApJ...865..142B}.
-This program will leverage the \tess\ Cycle 2 overlap with the \kepler\ prime field and portions of \ktwo\ fields.
+This program will leverage the significant overlap of the \tess\ Cycle 4 fields with the \ktwo\ fields.
 
@@
 For now we assume $I_\lambda(T_{spot})$ can be approximated as a cooler stellar photosphere, such as the \texttt{PHOENIX} model grid \cite{2013A&A...553A...6H}, an inexact but adequate assumption for the large bandpasses.} to $T_{spot}(\mathrm{SpT},\;P_{\mathrm{rot}})$.
+\newline \newline
+\textbf{Spot-induced modulation amplitudes depend on contrast.}
 
-\textbf{Spot-induced modulation amplitudes depend on contrast.}
 Rotationally-modulated lightcurves exhibit semi-sinusoidal undulations as starspots (or spot groups) enter and exit the projected stellar disk.
@@
 
-\textbf{Distinguishing Kepler/K2 and TESS amplitudes with statistically similar samples}
+\textbf{Head-to-head comparison of K2 and TESS amplitudes in the ecliptic fields}
 
-Ideally, we would have simultaneously observed the same targets with \kepler\ and \tess\ to measure the contrast for individual stars.
-Since this observational strategy was not possible for a large number of stars, we instead construct statistical subsamples of stars that are identical in every way, except for their observation band.
+Ideally, we would have simultaneously observed the same targets with \ktwo\ and \tess\ to measure the instantaneous amplitudes for individual stars.
+Cycle 4 provides a unique opportunity to compare the amplitudes of \tess\ and \ktwo\ for the largest sample to date.
 
@@
 
-\tess\ Cycle 2 will have significant overlap with both the \kepler\ prime field (up to 3 sectors of overlap) and 16438 source in the northern edges of \ktwo\ campaign fields (K.
-Col\'on, priv.
-comm.).
-This overlap sample offers an independent check on any statistical biases since the exact same stars can be compared.
-The fact that observations are not contemporaneous adds some \emph{cosmic variance}, but on average stars will enter and exit stellar activity cycles in equal measure, injecting per-star variance, but keeping the subsample mean amplitudes the same.
+\tess\ Cycle 4 will observe XX sources in the \ktwo\ campaign fields.
+The fact that observations are not contemporaneous adds some \emph{cosmic variance}: stars wander into and out-of their respective stellar minimum and maximum activity cycles.
+On average stars will enter and exit stellar activity cycles in equal measure, injecting per-star variance, but keeping the subsample mean amplitudes the same.
 
@@
 Measuring periods and amplitudes of \kepler\ and \tess\ lightcurves is a routine operation, with relatively low risk.
-The PI has significant experience determining amplitudes with a Fourier-based decomposition technique in young stars observed in \ktwo.
-The derived periods and amplitudes need not be \emph{perfect}, so long as the biases are about the same in both the \kepler\ and \tess\ subsamples.
 
 The yield of spotted stars in \tess\ will number in the hundreds of thousands.
-The \tess\ mission was designed to target cool main sequence stars with an abundance of K and M dwarfs.
+The \tess\ and \ktwo\ missions both target cool main sequence stars with an abundance of K and M dwarfs.
 This program will focus on the rapidly rotating stars ($P_{rot}<20$ days), which exhibit relatively large amplitudes ($A>2000$ ppm), illustrated in panel D of Figure \ref{fig:predictAmps}.
-This program will attempt---on a best effort basis---to measure the slowest rotating stars, which have periods much longer than the typical observation window of 27$-$54 days, and amplitudes comparable to the precision limits of \tess; these slow rotating sources must reside near the continuous viewing zone in the North Ecliptic Pole to have any hope of detecting spot modulation.
-
-One potential risk to this program is the that the \kepler\ and \tess\ populations will be statistically different from each other in their unobserved properties, confounding the assumption that the samples' only difference is the observation bandpass.
-For example, metallicity, age, extinction, and stellar background contamination may all be different between the \kepler\ samples and \tess\ samples.
-This risk is alleviated by the subsample of \kepler\ sources reobserved in \tess\ Cycle 2.
+This program will attempt---on a best effort basis---to measure slower rotating stars with ($P_{rot}<20$ days), which have periods comparable to typical observation windows of 27$-$54 days, and amplitudes comparable to the precision limits of \tess.
 
@@
 These probability distributions will empower exoplanet researchers to correct for biases in exoplanet transit radii.
-The probability distributions will reflect all the available information garnered from \tess, and reasonable assumptions about starspot geometrical distributions.
-
-Stellar surface structures serve as a stepping stone for brown dwarf cloud modeling, a poorly understood phenomenon that will inform our understanding of exoplanet spectroscopy.
-The proposed work will demonstrate how to gain insights about probable surface structures in the presence of significant geometrical degeneracies.
+The probability distributions will reflect all the available information garnered from \tess\ and reasonable assumptions about starspot geometrical distributions.
 
@@
 
-PI Gully-Santiago will carry out the data analysis using custom scripts that build upon the heritage of the \kepler\ data analysis ecosystem, including \texttt{lightkurve}, \texttt{astropy}, \texttt{celerite}, \texttt{BombScargle}, \texttt{PyTorch}, and \texttt{bokeh}.
-The PI will first apply these scripts to Kepler/K2 lightcurves while in the role as a support scientist at the Kepler/K2 Guest Observer Office.
-The PI will relocate to the University of Texas at Austin in February 2020 as a Research Fellow.
-UT Austin undergraduate and graduate students will lead the application of TESS-adapted scripts to \tess\ FFI lightcurves with \texttt{TESS-cut} and \texttt{elenor}.
+PI will carry out the data analysis using custom scripts that build upon the heritage of the \ktwo\ data analysis ecosystem, including \texttt{lightkurve} and \texttt{celerite}.
+The PI will first apply these scripts to \ktwo\ lightcurves.
+Undergraduate and graduate students will lead the application of TESS-adapted scripts to \tess\ FFI lightcurves with \texttt{TESS-cut} and \texttt{elenor}.
 Half of the funding will go to 1 undergraduate and 1 graduate student Summer 2020 salary, and 0.3 FTE to support the PI.
 All reproducible scripts---including Jupyter Notebook tutorials---will be publicly available and developed openly on GitHub.
-MGS will provide students with an interactive \texttt{bokeh} tool to visualize periodic lightcurve structures.
-This tool will leverage the PI's familiarity with \texttt{bokeh} to aid rapid visual spot-checking of the multidimensional analysis output.
+PI will provide students with an interactive tool to visualize periodic lightcurve structures to aid rapid visual spot-checking of the multidimensional analysis output.
 Finally, a High-Level Science Product (HLSP) will be delivered to MAST that contains the following information for each star found to possess detectable starspot-induced lightcurve modulation: 1) the period or periods, 2) an iteratively determined outlier mask of stellar flares, occultations, or spurious data artifacts, 3) a background-star amplitude-dilution estimate, 4) a probability distribution for the spot temperature and coverage fraction, 5) addition model fitting metadata.
-MGS will co-advise---with Co-I C.
-Morley---students in the writeup of the paper(s), authored and/or coauthored by students.
+PI will co-advise---with Co-I---students in the writeup of the paper(s), authored and/or coauthored by students.
 
@@
 
-\bibliography{tessgi_gully_cycle2}
+\bibliography{tessgi_gully_cycle4}
 

### 2021-01-22 proposals/TESS_GI_cycle04/tessgi_gully_cycle4.tex
Making headway on the TESS proposal for spot contrast

@@
 
-\usepackage{caption}
+\usepackage[svgnames]{xcolor}
+\definecolor{darkgrey}{RGB}{83, 92, 104}
+
+\usepackage[font={color=ocre,bf}]{caption}
 \usepackage{subcaption}
 \usepackage{cancel}
+\usepackage{wrapfig}
 
@@
 \newcommand{\kepler}{{\it Kepler}}
-\newcommand{\ktwo}{{K2}}
+\newcommand{\ktwo}{{\it K2}}
 \newcommand{\hst}{{\it HST}}
@@
 
-\section{Introduction}
-
 Starspots bias percieved exoplanet radii \cite{2018ApJ...853..122R}.
@@
 
-\section{Analysis Plan}
+\section{Analysis Plan: Compare \tess\ and \ktwo\ amplitudes}
 
@@
 
-Starspots emit light, which is often parameterized as \emph{spot contrast}, $c$, where 0 means the spots are non-emitting, and $c=1$ means the spot matches the specific intensity of the ambient photosphere.
-Spot contrast is wavelength dependent ($c_\lambda$) and can be thought of as possessing a characteristic spot temperature, $T_{spot}<T_{phot}$ that parameterizes the wavelength dependence of the contrasts.
-The analysis technique proposed here leverages the different wavelength dependence of the \kepler\ and \tess\ system throughputs to directly measure the ratio of the spot contrasts in their respective bandpasses.
-The spot contrast ratio---evaluated at these two pairs of bandpasses---provides a unique solution\footnote{Operationally, integrating the specific intensity $I_\lambda(T_{spot})$ over the system throughputs produces the expected contrast ratios.
-For now we assume $I_\lambda(T_{spot})$ can be approximated as a cooler stellar photosphere, such as the \texttt{PHOENIX} model grid \cite{2013A&A...553A...6H}, an inexact but adequate assumption for the large bandpasses.} to $T_{spot}(\mathrm{SpT},\;P_{\mathrm{rot}})$.
+Starspots emit light, which is often parameterized as \emph{spot contrast}, $c$, where $0$ means the spots are completely black compared to the bright photosphere, and $c=0.3$ means the spot has 30\% the flux per-unit-surface-area of the pristine photosphere.
+Hypothetically, a contrast of unity would mean the starspot blends into ambient photosphere, its border vanishing entirely.
+Spot contrast is wavelength dependent ($c_\lambda$).
+It has a spectrum of its own, starting close to zero in the blue/visible, and approaching perhaps 40\% in the near-IR; the exact prescription remains an active research area \cite{2005LRSP....2....8B}.
+Zeeman splitting and other magnetic phenomena are secondary to the mere fact that starspots appear as cooler-than-average regions on the stellar surface.
+
+Starspot flux can be parameterized by labeling the emission components with a characteristic temperatures $T_\text{spot}<T_\text{phot}$.
+The spot and ambient photosphere spectra depend only on those temperatures: $S(T_\text{spot})$ and $S(T_\text{phot})$.
+The wavelength-dependent spot contrast is then simply the ratio of the spot spectrum to the ambient photosphere spectrum:  $c_\lambda = \frac{S(T_\text{spot})}{ S(T_\text{phot})}$.
+
+Figure \ref{fig:filtercurve} shows an illustration of wavelength dependent spot contrast for a hypothetical M0V star with $T_\text{phot}=3800\;$K, and with $T_\text{spot}=3300\;$K.
+We employ \texttt{PHOENIX} synthetic spectral models \cite{2013A&A...553A...6H}, accessed at these temperatures.
+The contrast ascends from near 0.01 at $\lambda=500\;$nm, to 0.15 at 1 $\mu$m.
+By 2.3 $\mu$m, the contrast has peaked at 0.3.
+The overlapping \kepler\ and \tess\ system throughputs overplotted on Figure \ref{fig:filtercurve} indicate that---on average---the starspot contrast is a lower number for \kepler\ than for \tess.
+Starspots look blacker in \kepler\ and \ktwo\ than they do in TESS.
+The understanding of wavelength dependent starspot contrast is not novel--- it has driven the design of precision radial velocity (PRV) spectrographs to the near-IR, where the deleterious effect of starspots on RV jitter is lessened.
+\textbf{The key innovation of this proposal stems from the vast number of sources with existing \ktwo\ data that \tess\ will observe in Cycle 4.
+This proposal aims to quantify precisely the starspot contrast differences between the \tess\ and \kepler\ wavelength bands.}  We re-iterate that the extent to which nature produces starspot spectra resembling the prescription in Figure \ref{fig:filtercurve} remains an unknown, and is the main thrust of the proposed work effort.
+
+System-throughput-weighted contrast spectra can be integrated over the illustrated filter curves to yield exact estimates for contrasts in these two bands.
+For this M0V example, we obtain $c_{\ktwo}\sim0.02-0.3$ and $c_{\tess}\sim0.1$.
+Starspots appear about three to five times darker in \kepler\ and \ktwo\ than they do in \tess\.
+We define the ratio of starspot contrasts of \tess\ compared to \kepler\ as $\mathcal{R}\equiv \frac{c_{\tess}}{c_{\ktwo}}$.
+This proposal will measure the ratio of starspot contrast in the \tess\ bandpass compared to the \ktwo\ bandpass for a vast number of stars observed by both missions.
+
+illustrates the relative system throughputs of \tess\ and \kepler.
+
+The analysis technique proposed here will systematically measure the ratio of starspot contrast between $\lambda = 550$ nm and $850$ nm by taking advantage of the different wavelength dependence of the \kepler\ and \tess\ system throughputs.
+
+\captionsetup[figure]{font={color=darkgrey,md},textfont={small,singlespacing,md},labelfont={bf,normalsize,color=Black},labelformat={simple},labelsep=period,name={Figure}}
+
+\begin{figure}[hbt!]
+    \includegraphics[width=0.95\textwidth]{figures/contrast_spectrum.pdf}
+    \caption{Illustration of starspot contrast as seen in \tess\ and \kepler. }
+    \label{fig:filtercurve}
+\end{figure}
+
+The spot contrast ratio---evaluated at these two pairs of bandpasses---provides a unique solution Operationally, integrating the specific intensity $I_\lambda(T_{spot})$ over the system throughputs produces the expected contrast ratios.
+For now we assume $I_\lambda(T_{spot})$ can be approximated as a cooler stellar photosphere, such as the \texttt{PHOENIX} model grid \cite{2013A&A...553A...6H}, an inexact but adequate assumption for the large bandpasses.
+to $T_{spot}(\mathrm{SpT},\;P_{\mathrm{rot}})$.
 \newline \newline
@@
 
+\begin{wrapfigure}[23]{r}[10pt]{2.6in}
+    \vspace{0mm}
+    \begin{center}
+    \includegraphics[width=2.6in]{figures/fig1.pdf}
+    \caption{Expected signal strength of starspot contrast for the same stars measured in \tess\ and \kepler. }
+    \label{fig:expected}
+    \vspace{0mm}
+\end{center}
+\end{wrapfigure}
+
 Ideally, we would have simultaneously observed the same targets with \ktwo\ and \tess\ to measure the instantaneous amplitudes for individual stars.
@@
 
-\begin{figure}[hbt!]
-\includegraphics[width=0.95\textwidth]{figures/multi_panel.pdf}
-\caption{\texttt{A)} Schematic of a planet ($R_p = 0.1 R_\star$) transiting a star with unocculted starspots as the star moves from left to right.
-\texttt{B)} The resulting transit depth shows a systematic bias and a wavelength dependence due to starspots.
-\texttt{C)} Illustration of TLSE, \texttt{D)} Prediction for \tess\ amplitudes (colorful contours) relative to the \kepler\ sample (gray shaded background contours and black points) for $4000 < T_{phot} (K) < 4500$ \cite{2014ApJS..211...24M}.}
-\label{fig:predictAmps}
-\end{figure}
-
 \tess\ Cycle 4 will observe XX sources in the \ktwo\ campaign fields.
@@
 
-\section{Technical Feasibility}
+\section{Technical Feasibility: Consistency is key}
 

### 2021-01-22 proposals/TESS_GI_cycle04/tessgi_gully_cycle4.tex
More progress

@@
 
+\captionsetup[figure]{font={color=darkgrey,md},textfont={small,singlespacing,md},labelfont={bf,normalsize,color=Black},labelformat={simple},labelsep=period,name={Figure}}
+
+\begin{figure}[hbt!]
+    \includegraphics[width=0.95\textwidth]{figures/contrast_spectrum.pdf}
+    \caption{Illustration of starspot contrast as seen in \tess\ and \kepler.
+The red shaded spectrum illustrates the contrast for an M0V spectral type for a choice of spot and ambient photospheric temperatures.
+The thick blue left-most curve illustrates the \kepler\ passband, the thick green line to its right illustrates the \tess\ passband centered near 850 nm.
+A lower number means darker spots.
+}
+    \label{fig:filtercurve}
+\end{figure}
+
 Figure \ref{fig:filtercurve} shows an illustration of wavelength dependent spot contrast for a hypothetical M0V star with $T_\text{phot}=3800\;$K, and with $T_\text{spot}=3300\;$K.
@@
 By 2.3 $\mu$m, the contrast has peaked at 0.3.
-The overlapping \kepler\ and \tess\ system throughputs overplotted on Figure \ref{fig:filtercurve} indicate that---on average---the starspot contrast is a lower number for \kepler\ than for \tess.
-Starspots look blacker in \kepler\ and \ktwo\ than they do in TESS.
+The overlapping \kepler\ and \tess\ system throughputs overplotted on Figure \ref{fig:filtercurve} indicate that---on average---the starspot contrast is a lower number in the \kepler passband ($\lambda \sim 550\;$nm)  than for the redder \tess passband ($\lambda \sim 850\;$nm).
+Starspots look blacker in \kepler\ and \ktwo\ than they do in \tess.
 The understanding of wavelength dependent starspot contrast is not novel--- it has driven the design of precision radial velocity (PRV) spectrographs to the near-IR, where the deleterious effect of starspots on RV jitter is lessened.
@@
 
-System-throughput-weighted contrast spectra can be integrated over the illustrated filter curves to yield exact estimates for contrasts in these two bands.
-For this M0V example, we obtain $c_{\ktwo}\sim0.02-0.3$ and $c_{\tess}\sim0.1$.
-Starspots appear about three to five times darker in \kepler\ and \ktwo\ than they do in \tess\.
+System-throughput-weighted contrast spectra can be integrated over the illustrated filter curves to yield precision estimates for contrasts in the two passbands.
+For this M0V example, we obtain $c_{\ktwo}\sim0.025$ and $c_{\tess}\sim0.1$.
+Starspots appear about four times darker in \kepler\ and \ktwo\ than they do in \tess.
 We define the ratio of starspot contrasts of \tess\ compared to \kepler\ as $\mathcal{R}\equiv \frac{c_{\tess}}{c_{\ktwo}}$.
-This proposal will measure the ratio of starspot contrast in the \tess\ bandpass compared to the \ktwo\ bandpass for a vast number of stars observed by both missions.
+This proposal will systematically measure $\mathcal{R}$ for a vast number of stars observed by both missions.
 
-illustrates the relative system throughputs of \tess\ and \kepler.
-
-The analysis technique proposed here will systematically measure the ratio of starspot contrast between $\lambda = 550$ nm and $850$ nm by taking advantage of the different wavelength dependence of the \kepler\ and \tess\ system throughputs.
-
-\captionsetup[figure]{font={color=darkgrey,md},textfont={small,singlespacing,md},labelfont={bf,normalsize,color=Black},labelformat={simple},labelsep=period,name={Figure}}
-
-\begin{figure}[hbt!]
-    \includegraphics[width=0.95\textwidth]{figures/contrast_spectrum.pdf}
-    \caption{Illustration of starspot contrast as seen in \tess\ and \kepler. }
-    \label{fig:filtercurve}
-\end{figure}
-
-The spot contrast ratio---evaluated at these two pairs of bandpasses---provides a unique solution Operationally, integrating the specific intensity $I_\lambda(T_{spot})$ over the system throughputs produces the expected contrast ratios.
-For now we assume $I_\lambda(T_{spot})$ can be approximated as a cooler stellar photosphere, such as the \texttt{PHOENIX} model grid \cite{2013A&A...553A...6H}, an inexact but adequate assumption for the large bandpasses.
-to $T_{spot}(\mathrm{SpT},\;P_{\mathrm{rot}})$.
-\newline \newline
 \textbf{Spot-induced modulation amplitudes depend on contrast.}
-
 Rotationally-modulated lightcurves exhibit semi-sinusoidal undulations as starspots (or spot groups) enter and exit the projected stellar disk.

### 2021-01-22 proposals/TESS_GI_cycle04/tessgi_gully_cycle4.tex
Getting sorta mathy here, not sure that is a good thing

@@
 
-\textbf{Kepler and TESS perceive different starspot contrasts.}
+\noindent\textbf{Kepler and TESS perceive different starspot contrasts.}
 
@@
 This proposal will systematically measure $\mathcal{R}$ for a vast number of stars observed by both missions.
+\newline
 
-\textbf{Spot-induced modulation amplitudes depend on contrast.}
+\noindent \textbf{Spot-induced modulation amplitudes depend on contrast.}
+
 Rotationally-modulated lightcurves exhibit semi-sinusoidal undulations as starspots (or spot groups) enter and exit the projected stellar disk.
 The most-spotted projected hemisphere is directed towards the observer at the moment of the lightcurve minimum, and the least-spotted projected hemisphere at lightcurve maximum.
-The amplitude of the lightcurve, therefore, equals the difference in starspot coverage fractions $\Delta f_{spot} \equiv f_{max}-f_{min}$ between the most-and-least spotted projected hemispheres weighted by the spot contrast: $ A_\lambda = \Delta f_{spot} \cdot (1-c_\lambda)$.
+The amplitude of the lightcurve, therefore, is approximately the difference in starspot coverage fractions $\Delta f_{spot} \equiv f_{max}-f_{min}$ between the most-and-least spotted projected hemispheres, weighted by the spot contrast: $ A_\lambda = \Delta f_{spot} \cdot (1-c_\lambda)$.
 Therefore the ratio of \tess\ to \kepler\ amplitudes yields a constraint independent of the actual spot coverage fraction:
@@
 
-Recent stellar surface inhomogeneity modeling \cite{2018ApJ...853..122R, 2018ApJ...868..143G, 2018ApJ...865..142B} has yielded coarse scaling factors $\zeta$ for estimating the total starspot surface coverage fraction from the observed lightcurve amplitude $f_{spot}$, yielding $ p(f_{spot}) = \zeta \cdot A $, where $\zeta$ represents a probability distribution with values typically between 1 and 50, depending on the degree of assumed starspot longitudinal symmetry.
-Equipped solely with \tess-measured lightcurve amplitudes, we can provide what practitioners seek: a per-star estimate of $f_{spot}$ and $T_{spot}$.
+The notation is cumbersome, but the takeaway is clear: \textbf{we can quantify spot contrast ratio knowing only the amplitudes of \ktwo\ and \tess\ spot-modulated lightcurves.}  This equation is approximate, and in practice our team will derive $\mathcal{R}= \frac{c_{\tess}}{c_{\ktwo}}$ with Monte Carlo methods to take into account the uncertainty of the amplitude determination and subtle starspot self-dilution effects.
+\newline
 
-\textbf{Head-to-head comparison of K2 and TESS amplitudes in the ecliptic fields}
+\noindent \textbf{Head-to-head comparison of K2 and TESS amplitudes in the ecliptic fields}
 
@@
 The main two factors that affect spot-modulation lightcurve amplitudes are effective temperature (\emph{i.e.} mass), and rotation period (\emph{i.e.} stellar activity/magnetic fields) \cite{2014ApJS..211...24M}.
-Other important-albeit-unknown factors like stellar inclination angle and multiplicity are statistically identical between the \tess\ and \kepler\ samples.
+Other important-albeit-unknown factors like stellar inclination angle and multiplicity \emph{cancel out} when we compare the same stars head-to-head.
 We, therefore, propose to follow \cite{2014ApJS..211...24M} by constructing plots of amplitude versus period, subdivided into bins of effective temperature.
-The two samples should look identical in these plots, but the locus of \tess\ sources is expected to fall below the \kepler\ sample, as in panel D of Figure \ref{fig:predictAmps}.
+The two samples should look identical in these plots, but the locus of \tess\ sources is expected to fall below the \kepler\ sample, as in Figure \ref{fig:expected}.
 

### 2021-01-22 proposals/TESS_GI_cycle04/tessgi_gully_cycle4.tex
Done for the day!

@@
 
-Ideally, we would have simultaneously observed the same targets with \ktwo\ and \tess\ to measure the instantaneous amplitudes for individual stars.
-Cycle 4 provides a unique opportunity to compare the amplitudes of \tess\ and \ktwo\ for the largest sample to date.
-
 The main two factors that affect spot-modulation lightcurve amplitudes are effective temperature (\emph{i.e.} mass), and rotation period (\emph{i.e.} stellar activity/magnetic fields) \cite{2014ApJS..211...24M}.
 Other important-albeit-unknown factors like stellar inclination angle and multiplicity \emph{cancel out} when we compare the same stars head-to-head.
-We, therefore, propose to follow \cite{2014ApJS..211...24M} by constructing plots of amplitude versus period, subdivided into bins of effective temperature.
-The two samples should look identical in these plots, but the locus of \tess\ sources is expected to fall below the \kepler\ sample, as in Figure \ref{fig:expected}.
+We, therefore, construct a plot of amplitude versus period, grouped by spectral type bins.
+The \tess\ and \ktwo\ samples should look identical in these plots, but the locus of \tess\ sources resting systematically below the \ktwo\ sample, as in Figure \ref{fig:expected}.
 
-\tess\ Cycle 4 will observe XX sources in the \ktwo\ campaign fields.
-The fact that observations are not contemporaneous adds some \emph{cosmic variance}: stars wander into and out-of their respective stellar minimum and maximum activity cycles.
-On average stars will enter and exit stellar activity cycles in equal measure, injecting per-star variance, but keeping the subsample mean amplitudes the same.
+The fact that observations are not contemporaneous adds some \emph{cosmic variance}: stars wander into and out-of their respective stellar minimum and maximum activity cycles, altering their amplitudes in 2021 compared to say 2018.
+So it is therefore impossible to assign a starspot contrast ratio $\mathcal{R}$ for any individual star.
+Instead, we propose to measure the mean value of \emph{ensembles} of thousands of stars, again grouped by sources of similar properties.
+On average stars will enter and exit stellar activity cycles in equal measure injecting per-star variance, but keeping the subsample mean $\mathcal{R}$ the same.
 
-\section{Technical Feasibility: Consistency is key}
+\section{Technical Feasibility: Consistent and Systematic}
 
 Measuring periods and amplitudes of \kepler\ and \tess\ lightcurves is a routine operation, with relatively low risk.
+Amplitude estimates exist for some clusters, but with differing heuristics for for estimating the variability, such as interquartile range, sine-wave fitting, kernel methods, peak-to-valley, Fourier-based approaches, and other techniques that cannot be robustly intercompared to new \tess\ data.
+One centerpiece for this project is to compute modulation amplitudes systematically across both \ktwo\ and \tess.
 
-The yield of spotted stars in \tess\ will number in the hundreds of thousands.
-The \tess\ and \ktwo\ missions both target cool main sequence stars with an abundance of K and M dwarfs.
-This program will focus on the rapidly rotating stars ($P_{rot}<20$ days), which exhibit relatively large amplitudes ($A>2000$ ppm), illustrated in panel D of Figure \ref{fig:predictAmps}.
-This program will attempt---on a best effort basis---to measure slower rotating stars with ($P_{rot}<20$ days), which have periods comparable to typical observation windows of 27$-$54 days, and amplitudes comparable to the precision limits of \tess.
+The \ktwo\ mission targeted cool main sequence stars with an abundance of K and M dwarfs.
+This program will focus on the rapidly rotating stars ($P_{rot}<20$ days), which exhibit relatively large amplitudes ($A>2000$ ppm), illustrated in Figure \ref{fig:expected}.
+We estimate the total number of conceivable \tess\ and \ktwo\ rotationally modulated stars to be of order $10^5$, including cluster members, flare stars, active M-dwarfs, bright planet candidates, and other community targeted stellar programs acquired with long cadence postage stamps in \ktwo.
+This parent sample has to be vetted for \tess\ background overcrowding, \tess\ scattered light, variability classification provenance, signal-to-noise ratio, and other quality assurance checks.
+The mere curation of this sample represents a portion of the proposed work plan.
 
@@
 
-The main deliverable of this work will be a probabilistic assessment of the starspot temperature and starspot coverage faction for every star with a measured starspot modulation period and amplitude in \tess.
-These probability distributions will empower exoplanet researchers to correct for biases in exoplanet transit radii.
-The probability distributions will reflect all the available information garnered from \tess\ and reasonable assumptions about starspot geometrical distributions.
+The main deliverable of this work will be a probabilistic assessment of the starspot contrast ratio between \kepler\ and \tess\ as a function of spectral type and rotation rate.
+These probability distributions will empower exoplanet researchers to quantify biases in exoplanet transit radii, answering the question ``What starspot emission properties are plausible for my host-star of interest?''.
+For example, this work will make it possible to dispel certain categories of spot temperatures that could otherwise imbue spurious features in a JWST spectrum.
 
@@
 Trends in $T_{spot}$ with rotation rate and mass can reveal important clues to stellar activity, magnetic fields, and stellar atmospheres.
-This work will directly address starspot-induced scatter and bias in isochronal-derived stellar ages \cite{2015ApJ...807..174S, 2017ApJ...836..200G}.
+This work will inform starspot-induced scatter and bias in isochronal-derived stellar ages \cite{2015ApJ...807..174S, 2017ApJ...836..200G}, since spotted star SEDs can be tuned to fit the trends derived here.
 
@@
 
-PI will carry out the data analysis using custom scripts that build upon the heritage of the \ktwo\ data analysis ecosystem, including \texttt{lightkurve} and \texttt{celerite}.
-The PI will first apply these scripts to \ktwo\ lightcurves.
+PI will carry out the data analysis using a combination of new and existing scripts built upon the heritage of the \ktwo\ data analysis ecosystem, including \texttt{lightkurve} and \texttt{celerite}.
+The PI will first apply scripts to both \ktwo\ and \tess\ lightcurves.
 Undergraduate and graduate students will lead the application of TESS-adapted scripts to \tess\ FFI lightcurves with \texttt{TESS-cut} and \texttt{elenor}.
@@
 All reproducible scripts---including Jupyter Notebook tutorials---will be publicly available and developed openly on GitHub.
-PI will provide students with an interactive tool to visualize periodic lightcurve structures to aid rapid visual spot-checking of the multidimensional analysis output.
-Finally, a High-Level Science Product (HLSP) will be delivered to MAST that contains the following information for each star found to possess detectable starspot-induced lightcurve modulation: 1) the period or periods, 2) an iteratively determined outlier mask of stellar flares, occultations, or spurious data artifacts, 3) a background-star amplitude-dilution estimate, 4) a probability distribution for the spot temperature and coverage fraction, 5) addition model fitting metadata.
+PI will provide students with an interactive tool to visualize periodic lightcurve structures to aid rapid visual spot-checking of analysis output.
+Finally, a High-Level Science Product (HLSP) will be delivered to MAST that contains the following information for each star found to possess detectable starspot-induced lightcurve modulation: 1) the period or periods, 2) an iteratively determined outlier mask of stellar flares, occultations, or spurious data artifacts, 3) a background-star amplitude-dilution estimate, 4) a probability distribution for the spot temperature spot contrast ratio, 5) addition model fitting metadata.
 PI will co-advise---with Co-I---students in the writeup of the paper(s), authored and/or coauthored by students.

### 2022-06-01 papers/paper1/ms.tex
Add latex template and supporting file structure

\documentclass[modern]{aastex631}
\bibliographystyle{aasjournal}

% \usepackage{fontspec}
% \usepackage[T1]{fontenc}
% \usepackage{newtxsf}
% \setmainfont{Fira Sans Book}[Scale=1.0]

\usepackage[caption=false]{subfig}
\usepackage{booktabs}

\begin{document}
\shorttitle{Starspot contrast with TESS and K2}
\shortauthors{TBD}
\title{A systematic measurement on starspot contrast from TESS and K2 
bandpass differences}

\author{TBD}
\affiliation{University of Texas at Austin Department of Astronomy}

\author{TBD}
\affiliation{TBD}

\begin{abstract}

Abstract goes here.

\end{abstract}

\keywords{Starspots (1572)}

\section{Introduction}\label{sec:intro}

Here is an annotated bibliography.

\begin{deluxetable}{chc}
  \tablecaption{Annotated bibliography for intro\label{table1}}
  \tablehead{
  \colhead{Reference} & \nocolhead{two} & \colhead{Key idea}
  }
  \startdata
  \citet{gullysantiago17} & - & LkCa~4 starspot spectral decomposition\\
  \citet{2015ApJ...807..174S} & - & Radius inflation from spots \\
  Strassmeier & - & Starspots review \\
  Rackham & - & TLSE \\
  \enddata
\end{deluxetable}

\newpage

\begin{acknowledgements}
This paper includes data collected by the TESS mission. Funding for the TESS mission is provided by the NASA's Science Mission Directorate.

The authors acknowledge the Texas Advanced Computing Center (TACC, \url{http://www.tacc.utexas.edu}) at The University of Texas at Austin for providing HPC resources that have contributed to the research results reported within this paper.
\end{acknowledgements}

\clearpage

\facilities{Gemini:South, TESS, ASAS, Gaia}

\software{  pandas \citep{mckinney10, reback2020pandas},
  emcee \citep{foreman13},
  matplotlib \citep{hunter07},
  astroplan \citep{astroplan2018},
  astropy \citep{exoplanet:astropy13,exoplanet:astropy18},
  exoplanet \citep{exoplanet:exoplanet},
  numpy \citep{harris2020array},
  scipy \citep{jones01},
  ipython \citep{perez07},
  bokeh \citep{bokehcite},
  seaborn \citep{waskom14}}
  %pytorch \citep{NEURIPS2019_9015}} % No pytorch yet!

\bibliography{ms}

\clearpage

\appendix
\restartappendixnumbering

\section{Optional appendix} \label{appendix:tools}

Place optional content here.
\end{document}

### 2022-06-20 papers/paper1/ms.tex
Demo writing and start of real content to the paper
Cowriting the outline for the paper

@@
 \bibliographystyle{aasjournal}
+
+%\turnoffedit
 
@@
 \usepackage[caption=false]{subfig}
+\usepackage{censor}
 \usepackage{booktabs}
@@
 
-\newpage
+\section{Observations and Data Reduction}
+
+All sources were observed by both the K2 \citep{howell14} and TESS \citep{2015JATIS...1a4003R} missions.
+
+\subsection{Sample selection}
+Our initial sample mirrors...
+Hello!
+
+\subsection{TESS pipelines}
+
+\subsection{Quality assurance checks}
+
+\subsubsection{Handling data gaps}
+
+\subsubsection{Handling multiple sectors}
+
+\subsubsection{Quality flags}
+
+\subsubsection{Outlier detection}
+
+\authorcomment1{Outliers attributable to instrumental artifacts}.
+
+\authorcomment1{Outliers attributable to flares}.
+
+\authorcomment2{Outliers attributable to exoplanets}.
+
+\section{Analysis}
+
+\subsection{Period measurement}
+
+\subsubsection{Amplitude measurements}
+
+\section{Results}
+
+\subsection{Amplitude versus rotation period for bins in $T_{\mathrm{eff}}$}
+
+\subsection{TESS-to-K2 amplitude ratio as a function of $T_{\mathrm{eff}}$}
+
+\subsection{Converting amplitude ratio to $T_{\mathrm{spot}}$}
+
+\subsection{Inferred $T_{\mathrm{spot}}$ versus $T_{\mathrm{eff}}$}
+
+\section{Discussion}
+
+\subsection{Comparison to other studies}
+\citet{2016MNRAS.463.2494F} targeted 304 Pleiades sources with LAMOST and found the $T_{\mathrm{spot}}$ scales with $T_{\mathrm{eff}}$ with figure.
+
+\subsection{Assessing the spots versus faculae}
+Up to the this point, we have assumed that all flux variations arise from dark spots on the surface of the star.
+Stars can just as likely have bright patches that cause flux perturbations to their light curves.
+Here we assess the consequences of assuming faculae instead of spots.
+
+\section{Conclusions}
 

### 2022-07-15 papers/paper1/ms.tex
added some figures to the latex file
Started adding descriptions for the figures to the latex file

@@
 \usepackage{booktabs}
+\usepackage{graphicx}
+\graphicspath{{./figures/}}
 
@@
 Our initial sample mirrors...
-Hello!
 
@@
 
+Using the rapid rotator sample, we compared the amplitudes of the TESS and Kepler missions. We created a mask that removes outliers exceeding predetermined limitations, as seen below.
+
+  % Helpful resource on how to insert and align images
+  % https://www.overleaf.com/learn/latex/Inserting_Images
+  \begin{figure}[h]
+    \centering
+    \includegraphics[width=\linewidth]{Comparison Between TESS and Kepler Amplitudes.png}
+    \caption{Relationship between the calculated amplitudes of TESS and Kepler and how close they are to unity}
+  \end{figure}
+
 \section{Results}
@@
 \subsection{Amplitude versus rotation period for bins in $T_{\mathrm{eff}}$}
+
+Using the rapid rotator sample, we compared the rotational periods of both missions versus their amplitudes.
+This graph proves that our hypothesis from our proposal was correct.
+We estimate that the Kepler data is approximately 73 bigger that the TESS data (average amplitudes of 9353 versus 128, respectively).
+
+  \begin{figure}[h]
+    \centering
+    \includegraphics[width=\linewidth]{Amplitude vs. Rotation for Kepler and TESS.png}
+    \caption{Comparison between the rotational period vs. amplitude for both TESS and Kepler.}
+  \end{figure}
 

### 2022-07-16 papers/paper1/ms.tex
Working on pdf

--- 2022-06-20 papers/paper1/ms.tex
@@
 \usepackage{booktabs}
+\usepackage{graphicx}
+\usepackage{float}
+\usepackage[section]{placeins}
+\graphicspath{{./figures/}}
 
@@
 
+We compared the amplitudes of the TESS and Kepler spacecraft using the rapid rotator sample.
+As seen below, we developed a mask that eliminates outliers that beyond predefined boundaries.
+We decide to discard any points that exceed 80\% of the perfect correlation and points with over a 7 days period.
+The first requirement was decided because anything that differs by more than 80\%, in our opinion, most certainly the result of data fluctuations.
+Similarly, due to measurement errors, we have to exclude any periods longer than 7 days.
+The TESS mission only captured around 22 days worth of data for each object.
+We wanted to have at least 3 periods for each object.
+
+  % Helpful resource on how to insert and align images
+  % https://www.overleaf.com/learn/latex/Inserting_Images
+  \begin{figure}[!htb]
+    \centering
+    \includegraphics[scale=0.8]{Comparison Between TESS and Kepler Amplitudes.png}
+    \caption{Relationship between the calculated amplitudes of TESS and Kepler and how close they are to unity}
+  \end{figure}
+  \FloatBarrier
+
 \section{Results}
@@
 \subsection{TESS-to-K2 amplitude ratio as a function of $T_{\mathrm{eff}}$}
+
+ As seen in the figure below, the Kepler data are, on average, around 70 times larger than TESS's.
+
+  \begin{figure}[!htb]
+    \centering
+    \includegraphics[width=\linewidth]{Amplitude vs. Rotation for Kepler and TESS.png}
+    \caption{Comparison between the rotational period vs. amplitude for both TESS and Kepler. The Kepler and TESS averages are approximately 9350 and 130, respectively}
+  \end{figure}
+  \FloatBarrier
 
@@
 
-\software{  pandas \citep{mckinney10, reback2020pandas},
+\software{  
+  pandas \citep{mckinney10, reback2020pandas},
   emcee \citep{foreman13},

### 2022-07-18 papers/paper1/ms.tex
adding descriptions for figures in paper

@@
     \centering
-    \includegraphics[scale=0.8]{Comparison Between TESS and Kepler Amplitudes.png}
-    \caption{Relationship between the calculated amplitudes of TESS and Kepler and how close they are to unity}
+    \includegraphics[scale=0.5]{Comparison Between TESS and Kepler Amplitudes.png}
+    \caption{Relationship between the calculated amplitudes of TESS and Kepler and how close they are to unity.}
   \end{figure}
@@
     \centering
-    \includegraphics[width=\linewidth]{Amplitude vs. Rotation for Kepler and TESS.png}
-    \caption{Comparison between the rotational period vs. amplitude for both TESS and Kepler. The Kepler and TESS averages are approximately 9350 and 130, respectively}
+    \includegraphics[scale=0.42]{Amplitude vs. Rotation for Kepler and TESS.png}
+    \caption{Effect size for starspot contrast as seen in modulation amplitude versus period ($P_{\mathrm{rot}}$ [days]).
+The black points are the locus of spotted stars in the Kepler as measured by for late K dwarfs.
+The red points are the same stars, but measured with the TESS mission.
+The blue and green dashed lines represents the average amplitudes for all plotted stars for both Kepler and TESS, approximately 9350 and 130, respectively.}
   \end{figure}

### 2022-07-19 papers/paper1/ms.tex
Created new mask figure to show what each line means

@@
 \usepackage[section]{placeins}
+\usepackage{csquotes}
 \graphicspath{{./figures/}}
@@
 
+At first, the lightcurve program was falsely calculating many periods due to possible instrumental artifacts, flares, exoplanets, and background noise.
+To mitigate these outliers, we first tried using the \enquote{remove\_nans} and \enquote{remove\_outliers} methods.
+This helped, but still left many artifacts.
+Next, we decided to limit the period calculation to a maximum of 10 days and a minimum of 0.1 days.
+
+After we finished extrapolating the data using the computer, we then developed several criteria by close inspection to further eliminate outliers.
+One such criterion was to discard any points that exceeded or subceeded 80\% of perfect correlation between the missions.
+and points with over a 7 days period.
+The first requirement was decided because anything that differs by more than 80\%, in our opinion, most certainly the result of data fluctuations.
+Similarly, we have to exclude any periods greater than 7 days due to TESS's time restraints.
+TESS only collected data for each object for around 22 days, so we set a 7-day maximum period to include 3 whole wavelengths.
+
 \authorcomment1{Outliers attributable to instrumental artifacts}.
@@
 
+\begin{figure}[!htb]
+  \centering
+  \includegraphics[scale=0.5]{Amplitude Comparison Mask.png}
+  \caption{The mask showing how we determined which points were outliers.}
+\end{figure}
+\FloatBarrier
+
 \section{Analysis}
@@
 \subsubsection{Amplitude measurements}
-
-We compared the amplitudes of the TESS and Kepler spacecraft using the rapid rotator sample.
-As seen below, we developed a mask that eliminates outliers that beyond predefined boundaries.
-We decide to discard any points that exceed 80\% of the perfect correlation and points with over a 7 days period.
-The first requirement was decided because anything that differs by more than 80\%, in our opinion, most certainly the result of data fluctuations.
-Similarly, due to measurement errors, we have to exclude any periods longer than 7 days.
-The TESS mission only captured around 22 days worth of data for each object.
-We wanted to have at least 3 periods for each object.
 

### 2022-07-22 papers/paper1/ms.tex
Transfer some stock text from the proposal to the paper
subject to some rewriting yet to be done

@@
 \usepackage{graphicx}
+\usepackage{cancel}
 \usepackage{float}
@@
 \graphicspath{{./figures/}}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%% Some Useful Abbreviations %%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\newcommand{\tess}{{\it TESS}}
+\newcommand{\jwst}{{\it JWST}}
+\newcommand{\kepler}{{\it Kepler}}
+\newcommand{\ktwo}{{\it K2}}
+\newcommand{\hst}{{\it HST}}
+\newcommand{\msun}{$M_{\odot}$}
+\newcommand{\rsun}{$R_{\odot}$}
+\newcommand{\lsun}{$L_{\odot}$}
+\newcommand{\re}{$R_{\oplus}$}
+\newcommand{\me}{$M_{\oplus}$}
+\newcommand{\rj}{$R_{\textrm{\scriptsize Jup}}$}
+\newcommand{\mj}{$M_{\textrm{\scriptsize Jup}}$}
+\newcommand{\ms}{m~s$^{-1}$}
 
@@
 
-Here is an annotated bibliography.
+The diminution of flux arising from occultations of the stellar disk contains information the relative sizes of the occulting bodies.
+A commonsense exoplanet heuristic dictates that the relative transit depth $D$ equals the ratio of the bodies' projected areas ($R_p^2/R_\star^2$).
+Additional correction factors account for limb darkening \citep{2002ApJ...580L.171M}.
+Starspots on a stellar surface can exhibit spatial segregations
+that yield \emph{unocculted starspots}: spots that are present on the surface but are not traversed during transits.
+These unocculted starspots confound the mapping of transit depth to planet radius: a planet ``looks bigger'' than it really is if it blocks the brighter-on-average flux of the stellar disk \citep{2018AJ....156...91M}.
+You cannot measure a planet radius without making some assumptions about the extent, contrast, and distribution of starspots.
+
+The prevailing assumption within the exoplanet community has been that starspot covering fractions are small enough---as in Sunspots---to be ignored.
+New theoretical models \citep{2018ApJ...853..122R} and observational evidence \citep{2016MNRAS.463.2494F} indicate that starspot covering fractions are large enough to bias transit depths beyond the level of precision of \kepler\ and \tess.
+This so-called ``Transit Light Source Effect'' (TLSE) can bias derived exoplanet radii up to 17\% in the \tess\ bandpass, and can bias derived exoplanet \emph{densities} up to 50\%, since the radius error propagates into volume error with a factor of 3.
+Therefore, TLSE threatens placement and interpretation of the exoplanet mass-radius relation science requirements of the \tess\ mission.
+The interpretation of \jwst\ exoplanet transmission spectroscopy is also at risk \citep{2019AJ....157...11W}.
 
@@
 \section{Analysis}
+
+Starspots emit light, which is often parameterized as \emph{spot contrast}, $c$, where $0$ means the spots are completely black compared to the bright photosphere, and $c=0.3$ means the spot has 30\% the flux per-unit-surface-area of the pristine photosphere.
+Hypothetically, a contrast of unity would mean the starspot blends into ambient photosphere, its border vanishing entirely.
+Spot contrast is wavelength dependent ($c_\lambda$).
+It has a spectrum of its own, starting close to zero in the blue/visible, and approaching perhaps 40\% in the near-IR; the exact prescription remains an active research area \cite{2005LRSP....2....8B}.
+Zeeman splitting and other magnetic phenomena are secondary to the mere fact that starspots appear as cooler-than-average regions on the stellar surface.
+
+Starspot flux can be parameterized by labeling the emission components with a characteristic temperatures
+$T_\mathrm{spot}<T_\mathrm{phot}$.The spot and ambient photosphere spectra depend only on those temperatures: $S(T_\mathrm{spot})$ and $S(T_\mathrm{phot})$.
+The wavelength-dependent spot contrast is then simply the ratio of the spot spectrum to the ambient photosphere spectrum:  $c_\lambda = \frac{S(T_\mathrm{spot})}{ S(T_\mathrm{phot})}$.
+
+Figure \ref{fig:filtercurve} shows an illustration of wavelength dependent spot contrast for a hypothetical M0V star with $T_\mathrm{phot}=3800\;$K, and with $T_\mathrm{spot}=3300\;$K.
+We employ \texttt{PHOENIX} synthetic spectral models \cite{2013A&A...553A...6H}, accessed at these temperatures.
+The contrast ascends from near 0.01 at $\lambda=500\;$nm, to 0.15 at 1 $\mu$m.
+By 2.3 $\mu$m, the contrast has peaked at 0.3.
+The overlapping \kepler\ and \tess\ system throughputs overplotted on Figure \ref{fig:filtercurve} indicate that---on average---the starspot contrast is a lower number in the \kepler passband ($\lambda \sim 550\;$nm)  than for the redder \tess passband ($\lambda \sim 850\;$nm).
+Starspots look blacker in \kepler\ and \ktwo\ than they do in \tess.
+The understanding of wavelength dependent starspot contrast is not novel--- it has driven the design of precision radial velocity (PRV) spectrographs to the near-IR, where the deleterious effect of starspots on RV jitter is lessened.
+The key innovation of this proposal stems from the vast number of sources with existing \ktwo\ data that \tess\ will observe in Cycle 4.
+This proposal aims to quantify precisely the starspot contrast differences between the \tess\ and \kepler\ wavelength bands.
+We re-iterate that the extent to which nature produces starspot spectra resembling the prescription in Figure \ref{fig:filtercurve} remains an unknown, and is the main thrust of the proposed work effort.
+
+System-throughput-weighted contrast spectra can be integrated over the illustrated filter curves to yield precision estimates for contrasts in the two passbands.
+For this M0V example, we obtain $c_{\ktwo}\sim0.025$ and $c_{\tess}\sim0.1$.
+Starspots appear about four times darker in \kepler\ and \ktwo\ than they do in \tess.
+We define the ratio of starspot contrasts of \tess\ compared to \kepler\ as $\mathcal{R}\equiv \frac{c_{\tess}}{c_{\ktwo}}$.
+This proposal will systematically measure $\mathcal{R}$ for a vast number of stars observed by both missions.
+
+\noindent \textbf{Spot-induced modulation amplitudes depend on contrast.}
+
+Rotationally-modulated lightcurves exhibit semi-sinusoidal undulations as starspots (or spot groups) enter and exit the projected stellar disk.
+The most-spotted projected hemisphere is directed towards the observer at the moment of the lightcurve minimum, and the least-spotted projected hemisphere at lightcurve maximum.
+The amplitude of the lightcurve, therefore, is approximately the difference in starspot coverage fractions $\Delta f_{spot} \equiv f_{max}-f_{min}$ between the most-and-least spotted projected hemispheres, weighted by the spot contrast: $ A_\lambda = \Delta f_{spot} \cdot (1-c_\lambda)$.
+Therefore the ratio of \tess\ to \kepler\ amplitudes yields a constraint independent of the actual spot coverage fraction:
+
+$$ \frac{A_{TESS}}{A_{Kepler}} = \cancel{\frac{\Delta f_{spot}}{\Delta f_{spot}}} \cdot \frac{1-c_{TESS}}{1-c_{Kepler}} $$
+
+The notation is cumbersome, but the takeaway is clear: we can quantify spot contrast ratio knowing only the amplitudes of \ktwo\ and \tess\ spot-modulated lightcurves.
+This equation is approximate, and in practice our team will derive $\mathcal{R}= \frac{c_{\tess}}{c_{\ktwo}}$ with Monte Carlo methods to take into account the uncertainty of the amplitude determination and subtle starspot self-dilution effects.
+\newline
+
+\noindent \textbf{Head-to-head comparison of K2 and TESS amplitudes in the ecliptic fields}
+
+The main two factors that affect spot-modulation lightcurve amplitudes are effective temperature (\emph{i.e.} mass), and rotation period (\emph{i.e.} stellar activity/magnetic fields) \cite{2014ApJS..211...24M}.
+Other important-albeit-unknown factors like stellar inclination angle and multiplicity \emph{cancel out} when we compare the same stars head-to-head.
+We, therefore, construct a plot of amplitude versus period, grouped by spectral type bins.
+The \tess\ and \ktwo\ samples should look identical in these plots, but the locus of \tess\ sources resting systematically below the \ktwo\ sample, as in Figure \ref{fig:expected}.
+
+The fact that observations are not contemporaneous adds some \emph{cosmic variance}: stars wander into and out-of their respective stellar minimum and maximum activity cycles, altering their amplitudes in 2021 compared to say 2018.
+So it is therefore impossible to assign a starspot contrast ratio $\mathcal{R}$ for any individual star.
+Instead, we propose to measure the mean value of \emph{ensembles} of thousands of stars, again grouped by sources of similar properties.
+On average stars will enter and exit stellar activity cycles in equal measure injecting per-star variance, but keeping the subsample mean $\mathcal{R}$ the same.
 

### 2022-07-23 papers/paper1/ms.tex
Add figure in latex, but not yet committed file

@@
 
+\begin{figure}[hbt!]
+  \includegraphics[width=0.95\textwidth]{figures/contrast_spectrum2.pdf}
+  \caption{Illustration of starspot contrast as seen in \tess\ and \kepler.
+The red shaded spectrum illustrates the contrast for an M0V spectral type for a choice of spot and ambient photospheric temperatures.
+The thick blue left-most curve illustrates the \kepler\ passband, the thick green line to its right illustrates the \tess\ passband centered near 850 nm.
+A lower number means darker spots.
+}
+  \label{fig:filtercurve}
+\end{figure}
+
 Figure \ref{fig:filtercurve} shows an illustration of wavelength dependent spot contrast for a hypothetical M0V star with $T_\mathrm{phot}=3800\;$K, and with $T_\mathrm{spot}=3300\;$K.
@@
 We, therefore, construct a plot of amplitude versus period, grouped by spectral type bins.
-The \tess\ and \ktwo\ samples should look identical in these plots, but the locus of \tess\ sources resting systematically below the \ktwo\ sample, as in Figure \ref{fig:expected}.
+The \tess\ and \ktwo\ samples should look identical in these plots, but the locus of \tess\ sources resting systematically below the \ktwo\ sample.
 

### 2022-07-25 papers/paper1/ms.tex
Reran the data with normalized lightcurves and recreated the figures

--- 2022-07-22 papers/paper1/ms.tex
@@
   \includegraphics[scale=0.5]{Amplitude Comparison Mask.png}
-  \caption{The mask showing how we determined which points were outliers.}
+  \caption{The mask showing how we determined which points to keep and which ones were outliers.}
 \end{figure}
@@
 The wavelength-dependent spot contrast is then simply the ratio of the spot spectrum to the ambient photosphere spectrum:  $c_\lambda = \frac{S(T_\mathrm{spot})}{ S(T_\mathrm{phot})}$.
+
+% \begin{figure}[hbt!]
+%   \includegraphics[width=0.95\textwidth]{figures/contrast_spectrum2.pdf}
+%   \caption{Illustration of starspot contrast as seen in \tess\ and \kepler.
+The red shaded spectrum illustrates the contrast for an M0V spectral type for a choice of spot and ambient photospheric temperatures.
+The thick blue left-most curve illustrates the \kepler\ passband, the thick green line to its right illustrates the \tess\ passband centered near 850 nm.
+A lower number means darker spots.}
+%   \label{fig:filtercurve}
+% \end{figure}
 
@@
 
-As seen in the figure below, the Kepler data are, on average, around 70 times larger than TESS's.
+As seen in the figure below, the TESS data are, on average, around 175\% larger than Kepler.
 
@@
 The red points are the same stars, but measured with the TESS mission.
-The blue and green dashed lines represents the average amplitudes for all plotted stars for both Kepler and TESS, approximately 9350 and 130, respectively.}
+The blue and green dashed lines represents the average amplitudes for all plotted stars for both Kepler and TESS, approximately 3.98\% and 6.94\%, respectively.}
 \end{figure}

### 2022-07-28 papers/paper1/ms.tex
Added some more citations

@@
 \graphicspath{{./figures/}}
+
+% Enable for dark mode
+% \usepackage{xcolor}
+% \pagecolor{black}
+% \color{white}
 
@@
   Strassmeier & - & Starspots review \\
-  Rackham & - & TLSE \\
+  \citet{2018ApJ...853..122R} & - & TLSE \\
+  \citet{Kochukhov21} & - & Magnetic fields of M dwarfs \\
+  \citet{2021sf2a.conf..363B} & - & Stellar rotation using ROOSTER \\
+  \citet{2015MNRAS.450.3211A} & - & Using Kepler to study stellar rotation \\
+  \citet{2022ApJ...927..219C} & - & Using deep learning to recover TESS light curves \\
+  \citet{2021AJ....162..123L} & - & Test Lightcurve and study stellar surfaces \\
+  \citet{2022ApJ...925..185D} & - & Lightcurves for oblate, rapid rotators \\
+  \citet{Reinhold2013} & - & Measure rotation of active Kepler Stars \\
+  % \citet{} & - & \\
+  % \citet{} & - & \\
+  % \citet{} & - & \\
+  % \citet{} & - & \\
+  % \citet{} & - & \\
+  % \citet{} & - & \\
+  % \citet{} & - & \\
+  % \citet{} & - & \\  
   \enddata

### 2022-07-29 papers/paper1/ms.tex

@@
 
-% \begin{figure}[hbt!]
-%   \includegraphics[width=0.95\textwidth]{figures/contrast_spectrum2.pdf}
-%   \caption{Illustration of starspot contrast as seen in \tess\ and \kepler.
+\begin{figure}[hbt!]
+  \includegraphics[width=0.95\textwidth]{figures/contrast_spectrum2.pdf}
+  \caption{Illustration of starspot contrast as seen in \tess\ and \kepler.
 The red shaded spectrum illustrates the contrast for an M0V spectral type for a choice of spot and ambient photospheric temperatures.
 The thick blue left-most curve illustrates the \kepler\ passband, the thick green line to its right illustrates the \tess\ passband centered near 850 nm.
-A lower number means darker spots.}
-%   \label{fig:filtercurve}
-% \end{figure}
+A lower number means darker spots.
+}
+  \label{fig:filtercurve}
+\end{figure}
 

### 2022-07-29 papers/paper1/ms.tex
Add figures of contrast versus Teff ratio

@@
-\documentclass[modern]{aastex631}
+\documentclass[twocolumn]{aastex631}
 \bibliographystyle{aasjournal}
@@
 
-\begin{deluxetable}{chc}
+\begin{deluxetable*}{chc}
   \tablecaption{Annotated bibliography for intro\label{table1}}
@@
   \enddata
-\end{deluxetable}
+\end{deluxetable*}
 
@@
 
-\begin{figure}[!htb]
+\begin{figure*}[!htb]
   \centering
@@
   \caption{The mask showing how we determined which points to keep and which ones were outliers.}
-\end{figure}
+\end{figure*}
 \FloatBarrier
@@
 
-\begin{figure}[hbt!]
+\begin{figure*}[hbt!]
   \includegraphics[width=0.95\textwidth]{figures/contrast_spectrum2.pdf}
@@
   \label{fig:filtercurve}
-\end{figure}
+\end{figure*}
 
-Figure \ref{fig:filtercurve} shows an illustration of wavelength dependent spot contrast for a hypothetical M0V star with $T_\mathrm{phot}=3800\;$K, and with $T_\mathrm{spot}=3300\;$K.
+Figure \ref{fig:filtercurve} shows an illustration of wavelength dependent spot contrast for a hypothetical K star with $T_\mathrm{phot}=4500\;$K, and with $T_\mathrm{spot}=2700\;$K.
 We employ \texttt{PHOENIX} synthetic spectral models \cite{2013A&A...553A...6H}, accessed at these temperatures.
@@
 We define the ratio of starspot contrasts of \tess\ compared to \kepler\ as $\mathcal{R}\equiv \frac{c_{\tess}}{c_{\ktwo}}$.
-This proposal will systematically measure $\mathcal{R}$ for a vast number of stars observed by both missions.
 
@@
 
-The notation is cumbersome, but the takeaway is clear: we can quantify spot contrast ratio knowing only the amplitudes of \ktwo\ and \tess\ spot-modulated lightcurves.
+For our example with a 4500 K ambient photosphere and 2700 K starspot characteristic temperature, we obtained $c_{\ktwo}\sim0.025$ and $c_{\tess}\sim0.1$, yielding an expected amplitude ratio of:
+
+$$\frac{1-c_{TESS}}{1-c_{Kepler}} = \frac{1-0.1}{1-0.025}  = (0.9/0.975)=0.923$$
+
+\authorcomment1{We are not currently weighting by the TESS and K2 filter curves.  Here are the values we get when we take the mean:}
+
+$$\frac{1-c_{TESS}}{1-c_{Kepler}} = \frac{1-0.104}{1-0.00697}  = (0.896/0.99303)=0.902$$
+
+% TESS - 0.104
+% Kepler - 0.00697
+
 This equation is approximate, and in practice our team will derive $\mathcal{R}= \frac{c_{\tess}}{c_{\ktwo}}$ with Monte Carlo methods to take into account the uncertainty of the amplitude determination and subtle starspot self-dilution effects.
@@
 
+\subsection{Starspot contrast theory}
+Tbd...
+
+\begin{figure}
+  \gridline{\fig{figures/Kepler_contrast.png}{0.3\textwidth}{(a)}}
+  \gridline{\fig{figures/TESS_contrast.png}{0.3\textwidth}{(b)}}
+  \gridline{\fig{one_minus_ratio.png}{0.3\textwidth}{(c)}}
+  \caption{A nice inverted pyramid figure
+    consisting of six individual files}
+\end{figure}
+
 \section{Results}
@@
   \centering
-  \includegraphics[scale=0.42]{Amplitude vs. Rotation for Kepler and TESS.png}
+  \includegraphics[width=0.95\columnwidth]{Amplitude vs. Rotation for Kepler and TESS.png}
   \caption{Effect size for starspot contrast as seen in modulation amplitude versus period ($P_{\mathrm{rot}}$ [days]).

### 2022-07-31 papers/paper1/ms.tex
Moved CSV files, created new notebook, and added more citations to paper

@@
   \startdata
-  \citet{gullysantiago17} & - & LkCa~4 starspot spectral decomposition\\
+  \citet{2015MNRAS.450.3211A} & - & Using Kepler to study stellar rotation \\
+  \citet{2022ApJ...924...31B} & - & Estimating the lifetime of starspots based on rotational periods \\
+  \citet{2010ApJ...713L.109B} & - & Why Kepler picked particular stars for observation \\
+  \citet{2010Sci...327..977B} & - & Using the Kepler mission to discover exoplanets \\
+  \citet{2021sf2a.conf..363B} & - & Stellar rotation using ROOSTER \\
+  \citet{2022ApJ...927..219C} & - & Deep learning to recover TESS light curves \\
+  \citet{2022ApJ...925..185D} & - & Lightcurves for oblate, rapid rotators \\
+  \citet{2017ARep...61..122D} & - & Using Kepler to observe starspots on M dwarf stars \\
+  \citet{2020RAA....20..167F} & - & Overview of Kepler mission \\
+  \citet{gullysantiago17} & - & LkCa~4 starspot spectral decomposition \\
+  \citet{Kochukhov21} & - & Magnetic fields of M dwarfs \\
+  \citet{2021AJ....162..123L} & - & Test Lightcurve and study stellar surfaces \\
+  \citet{2022ApJ...933..195M} & - & Comparing rotational periods of sun-like stars versus photometric amplitude \\
+  \citet{2020MNRAS.491.2706N} & - & Calculating stellar age based on starspot covering fraction \\
+  \citet{2018ApJ...853..122R} & - & TLSE \\
+  \citet{Reinhold2013} & - & Measure rotation of active Kepler Stars \\
   \citet{2015ApJ...807..174S} & - & Radius inflation from spots \\
   Strassmeier & - & Starspots review \\
-  \citet{2018ApJ...853..122R} & - & TLSE \\
-  \citet{Kochukhov21} & - & Magnetic fields of M dwarfs \\
-  \citet{2021sf2a.conf..363B} & - & Stellar rotation using ROOSTER \\
-  \citet{2015MNRAS.450.3211A} & - & Using Kepler to study stellar rotation \\
-  \citet{2022ApJ...927..219C} & - & Using deep learning to recover TESS light curves \\
-  \citet{2021AJ....162..123L} & - & Test Lightcurve and study stellar surfaces \\
-  \citet{2022ApJ...925..185D} & - & Lightcurves for oblate, rapid rotators \\
-  \citet{Reinhold2013} & - & Measure rotation of active Kepler Stars \\
-  % \citet{} & - & \\
-  % \citet{} & - & \\
-  % \citet{} & - & \\
-  % \citet{} & - & \\
-  % \citet{} & - & \\
-  % \citet{} & - & \\
-  % \citet{} & - & \\
-  % \citet{} & - & \\  
+  \citet{2022MNRAS.511.2285W} & - & Starspots on detached eclipsing binaries (DEBs) \\
   \enddata
@@
 
-\begin{figure*}[!htb]
+\begin{figure}[!htb]
   \centering
-  \includegraphics[scale=0.5]{Amplitude Comparison Mask.png}
+  \includegraphics[width=0.5\textwidth]{Amplitude Comparison Mask.png}
   \caption{The mask showing how we determined which points to keep and which ones were outliers.}
-\end{figure*}
+\end{figure}
 \FloatBarrier
@@
 
-% TESS - 0.104
-% Kepler - 0.00697
-
 This equation is approximate, and in practice our team will derive $\mathcal{R}= \frac{c_{\tess}}{c_{\ktwo}}$ with Monte Carlo methods to take into account the uncertainty of the amplitude determination and subtle starspot self-dilution effects.
@@
 
-\begin{figure}
-  \gridline{\fig{figures/Kepler_contrast.png}{0.3\textwidth}{(a)}}
-  \gridline{\fig{figures/TESS_contrast.png}{0.3\textwidth}{(b)}}
-  \gridline{\fig{one_minus_ratio.png}{0.3\textwidth}{(c)}}
+\begin{figure*}
+  \gridline
+    {\fig{figures/Kepler_contrast.png}{0.3\textwidth}{(a)}
+    {\fig{figures/TESS_contrast.png}{0.3\textwidth}{(b)}}
+    {\fig{one_minus_ratio.png}{0.3\textwidth}{(c)}}}
   \caption{A nice inverted pyramid figure
     consisting of six individual files}
-\end{figure}
+\end{figure*}
 

### 2022-08-01 papers/paper1/ms.tex
Added equation from Rackham
Writing prose-- some introduction, some methods, some equation placeholders.

@@
 
+\authorcomment1{General background information and motivation for starspots...}
+
+The simplest observational predictors of starspot physical properties appear to be effective temperature (\emph{i.e} spectral type, mass, and/or evolutionary state) and rotation rate (\emph{i.e.} Rossby number).
+Theese terms apparently control the conditions that give rise to spots \censorbox{Cite Mayukh Panja papers}.
+
+Disentangling starspot contrast and area is challenging, since many observables produce degeneracies among these two distinct physical properties.
+For example, a given flux loss in a monochromatic time series lightcurve can indicate a large warm starspot or or a small cool starspot.
+
 The diminution of flux arising from occultations of the stellar disk contains information the relative sizes of the occulting bodies.
@@
   \citet{2022MNRAS.511.2285W} & - & Starspots on detached eclipsing binaries (DEBs) \\
+  Mayukh Panja & What do starspots look like \\
   \enddata
 \end{deluxetable*}
+
+\section{Methodology}
+
+\subsection{Starspot contrast and temperature}
+There does not appear to be a first-principles theoretical prediction for the spectrum of a starspot, as a function of the physical state of the star.
+Instead, practitioners have come up with phenomenological models based on observed trends in spot contrast coupled with ostensibly commonsense heuristics.
+Most practitioners assume that starspot temperature depends principally on the photospheric temperature.
+Two sources provide similar but slightly different trends for this relationship.
+\citet{2019AJ....157...96R} distilled the compilation of spot temperatures from \citet{2005LRSP....2....8B} into a simple first order linear relationship:
+\begin{equation}
+  T_{\mathrm{spot}} = 0.418 \times T_{\mathrm{phot}} + 1620~\mathrm{K}
+  \label{eq:Tspot}
+\end{equation}
+
+\citet{2016MNRAS.463.2494F} report an ostensibly quadratic dependence on the ambient photosphere, with hot stars asympoting to a typical starspot temperature of $\sim3500~\mathrm{K}$.
+This departure from the linear trend could encode an observational bias since the LAMOST TiO-based technique may lack sensitivity to hypotetical hotter spots.
+Nevertheless the trend serves as a useful guide in the absence of more, and higher quality, data.
+Here we fit a quadratic relationship to the \citet{2016MNRAS.463.2494F} LAMOST-derived temperature points to obtain the relationship:
+\authorcomment1{Ryan H. to determine A, B, and C}.
+
+\begin{equation}
+  T_{\mathrm{spot}} = a \times T_{\mathrm{phot}}^2 + b \times T_{\mathrm{phot}} + c ~\mathrm{K}
+  \label{eq:TspotFang}
+\end{equation}
+
+\authorcomment1{Todo: Cite Lyra Cao et al. in prep-- estimated August 2022}.
+
+We plot both trends in Figure \censorbox{X}.  The range of validity and accuracy of these relationships remain unclear. Interrogating these trends forms the main thrust of this work.
+
+\subsection{Starspot contrast from Kepler and TESS}
+A typical starspot
 

### 2022-08-01 papers/paper1/ms.tex
Rearrange and rewrite some text

@@
 
-\subsection{Starspot contrast and temperature}
-There does not appear to be a first-principles theoretical prediction for the spectrum of a starspot, as a function of the physical state of the star.
+\subsection{Starspot contrast}
+Starspots emit light, which is often parameterized as spot contrast, $c$, where $0$ means the spots are completely black compared to the bright photosphere, and $c=0.3$ means the spot has 30\% the flux per-unit-surface-area of the pristine photosphere.
+Hypothetically, a contrast of unity would mean the starspot blends into ambient photosphere, its border vanishing entirely.
+Spot contrast is wavelength dependent ($c_\lambda$).
+It has a spectrum of its own, starting close to zero in the blue/visible, and approaching perhaps 40\% in the near-IR.
+
+There does not appear to be a first-principles theoretical prediction for the spectrum of a starspot as a function of the fundamental properties of the star.
 Instead, practitioners have come up with phenomenological models based on observed trends in spot contrast coupled with ostensibly commonsense heuristics.
-Most practitioners assume that starspot temperature depends principally on the photospheric temperature.
+For example, most practitioners assume that starspot temperature depends principally on the photospheric temperature.
 Two sources provide similar but slightly different trends for this relationship.
@@
 \citet{2016MNRAS.463.2494F} report an ostensibly quadratic dependence on the ambient photosphere, with hot stars asympoting to a typical starspot temperature of $\sim3500~\mathrm{K}$.
-This departure from the linear trend could encode an observational bias since the LAMOST TiO-based technique may lack sensitivity to hypotetical hotter spots.
+This departure from the linear trend could be real, or may encode an observational bias since the LAMOST TiO-based technique may lack sensitivity to hypotetical hotter spots.
 Nevertheless the trend serves as a useful guide in the absence of more, and higher quality, data.
@@
 We plot both trends in Figure \censorbox{X}.  The range of validity and accuracy of these relationships remain unclear. Interrogating these trends forms the main thrust of this work.
+
+Starspot flux can be parameterized by labeling the emission components with a characteristic temperatures
+$T_\mathrm{spot}<T_\mathrm{phot}$.The spot and ambient photosphere spectra depend only on those temperatures: $S(T_\mathrm{spot})$ and $S(T_\mathrm{phot})$.
+The wavelength-dependent spot contrast is then simply the ratio of the spot spectrum to the ambient photosphere spectrum:  $c_\lambda = \frac{S(T_\mathrm{spot})}{ S(T_\mathrm{phot})}$.
+
+\begin{figure*}[hbt!]
+  \includegraphics[width=0.95\textwidth]{figures/contrast_spectrum2.pdf}
+  \caption{Illustration of starspot contrast as seen in \tess\ and \kepler.
+The red shaded spectrum illustrates the contrast for an M0V spectral type for a choice of spot and ambient photospheric temperatures.
+The thick blue left-most curve illustrates the \kepler\ passband, the thick green line to its right illustrates the \tess\ passband centered near 850 nm.
+A lower number means darker spots.
+}
+  \label{fig:filtercurve}
+\end{figure*}
+
+Figure \ref{fig:filtercurve} shows an illustration of wavelength dependent spot contrast for a hypothetical K star with $T_\mathrm{phot}=4500\;$K, and with $T_\mathrm{spot}=2700\;$K.
+We employ \texttt{PHOENIX} synthetic spectral models \cite{2013A&A...553A...6H}, accessed at these temperatures.
+The contrast ascends from near 0.01 at $\lambda=500\;$nm, to 0.15 at 1 $\mu$m.
+By 2.3 $\mu$m, the contrast has peaked at 0.3.
+The overlapping \kepler\ and \tess\ system throughputs overplotted on Figure \ref{fig:filtercurve} indicate that---on average---the starspot contrast is a lower number in the \kepler passband ($\lambda \sim 550\;$nm)  than for the redder \tess passband ($\lambda \sim 850\;$nm).
+Starspots look blacker in \kepler\ and \ktwo\ than they do in \tess.
+The understanding of wavelength dependent starspot contrast is not novel--- it has driven the design of precision radial velocity (PRV) spectrographs to the near-IR, where the deleterious effect of starspots on RV jitter is lessened.
+The key innovation of this proposal stems from the vast number of sources with existing \ktwo\ data that \tess\ will observe in Cycle 4.
+This proposal aims to quantify precisely the starspot contrast differences between the \tess\ and \kepler\ wavelength bands.
+We re-iterate that the extent to which nature produces starspot spectra resembling the prescription in Figure \ref{fig:filtercurve} remains an unknown, and is the main thrust of the proposed work effort.
+
+System-throughput-weighted contrast spectra can be integrated over the illustrated filter curves to yield precision estimates for contrasts in the two passbands.
+For this M0V example, we obtain $c_{\ktwo}\sim0.025$ and $c_{\tess}\sim0.1$.
+Starspots appear about four times darker in \kepler\ and \ktwo\ than they do in \tess.
+We define the ratio of starspot contrasts of \tess\ compared to \kepler\ as $\mathcal{R}\equiv \frac{c_{\tess}}{c_{\ktwo}}$.
+
+\noindent \textbf{Spot-induced modulation amplitudes depend on contrast.}
+
+Rotationally-modulated lightcurves exhibit semi-sinusoidal undulations as starspots (or spot groups) enter and exit the projected stellar disk.
+The most-spotted projected hemisphere is directed towards the observer at the moment of the lightcurve minimum, and the least-spotted projected hemisphere at lightcurve maximum.
+The amplitude of the lightcurve, therefore, is approximately the difference in starspot coverage fractions $\Delta f_{spot} \equiv f_{max}-f_{min}$ between the most-and-least spotted projected hemispheres, weighted by the spot contrast: $ A_\lambda = \Delta f_{spot} \cdot (1-c_\lambda)$.
+Therefore the ratio of \tess\ to \kepler\ amplitudes yields a constraint independent of the actual spot coverage fraction:
+
+$$ \frac{A_{TESS}}{A_{Kepler}} = \cancel{\frac{\Delta f_{spot}}{\Delta f_{spot}}} \cdot \frac{1-c_{TESS}}{1-c_{Kepler}} $$
+
+For our example with a 4500 K ambient photosphere and 2700 K starspot characteristic temperature, we obtained $c_{\ktwo}\sim0.025$ and $c_{\tess}\sim0.1$, yielding an expected amplitude ratio of:
+
+$$\frac{1-c_{TESS}}{1-c_{Kepler}} = \frac{1-0.1}{1-0.025}  = (0.9/0.975)=0.923$$
+
+\authorcomment1{We are not currently weighting by the TESS and K2 filter curves.  Here are the values we get when we take the mean:}
+
+$$\frac{1-c_{TESS}}{1-c_{Kepler}} = \frac{1-0.104}{1-0.00697}  = (0.896/0.99303)=0.902$$
+
+This equation is approximate, and in practice our team will derive $\mathcal{R}= \frac{c_{\tess}}{c_{\ktwo}}$ with Monte Carlo methods to take into account the uncertainty of the amplitude determination and subtle starspot self-dilution effects.
 
@@
 \section{Analysis}
-
-Starspots emit light, which is often parameterized as \emph{spot contrast}, $c$, where $0$ means the spots are completely black compared to the bright photosphere, and $c=0.3$ means the spot has 30\% the flux per-unit-surface-area of the pristine photosphere.
-Hypothetically, a contrast of unity would mean the starspot blends into ambient photosphere, its border vanishing entirely.
-Spot contrast is wavelength dependent ($c_\lambda$).
-It has a spectrum of its own, starting close to zero in the blue/visible, and approaching perhaps 40\% in the near-IR; the exact prescription remains an active research area \cite{2005LRSP....2....8B}.
-Zeeman splitting and other magnetic phenomena are secondary to the mere fact that starspots appear as cooler-than-average regions on the stellar surface.
-
-Starspot flux can be parameterized by labeling the emission components with a characteristic temperatures
-$T_\mathrm{spot}<T_\mathrm{phot}$.The spot and ambient photosphere spectra depend only on those temperatures: $S(T_\mathrm{spot})$ and $S(T_\mathrm{phot})$.
-The wavelength-dependent spot contrast is then simply the ratio of the spot spectrum to the ambient photosphere spectrum:  $c_\lambda = \frac{S(T_\mathrm{spot})}{ S(T_\mathrm{phot})}$.
-
-\begin{figure*}[hbt!]
-  \includegraphics[width=0.95\textwidth]{figures/contrast_spectrum2.pdf}
-  \caption{Illustration of starspot contrast as seen in \tess\ and \kepler.
-The red shaded spectrum illustrates the contrast for an M0V spectral type for a choice of spot and ambient photospheric temperatures.
-The thick blue left-most curve illustrates the \kepler\ passband, the thick green line to its right illustrates the \tess\ passband centered near 850 nm.
-A lower number means darker spots.
-}
-  \label{fig:filtercurve}
-\end{figure*}
-
-Figure \ref{fig:filtercurve} shows an illustration of wavelength dependent spot contrast for a hypothetical K star with $T_\mathrm{phot}=4500\;$K, and with $T_\mathrm{spot}=2700\;$K.
-We employ \texttt{PHOENIX} synthetic spectral models \cite{2013A&A...553A...6H}, accessed at these temperatures.
-The contrast ascends from near 0.01 at $\lambda=500\;$nm, to 0.15 at 1 $\mu$m.
-By 2.3 $\mu$m, the contrast has peaked at 0.3.
-The overlapping \kepler\ and \tess\ system throughputs overplotted on Figure \ref{fig:filtercurve} indicate that---on average---the starspot contrast is a lower number in the \kepler passband ($\lambda \sim 550\;$nm)  than for the redder \tess passband ($\lambda \sim 850\;$nm).
-Starspots look blacker in \kepler\ and \ktwo\ than they do in \tess.
-The understanding of wavelength dependent starspot contrast is not novel--- it has driven the design of precision radial velocity (PRV) spectrographs to the near-IR, where the deleterious effect of starspots on RV jitter is lessened.
-The key innovation of this proposal stems from the vast number of sources with existing \ktwo\ data that \tess\ will observe in Cycle 4.
-This proposal aims to quantify precisely the starspot contrast differences between the \tess\ and \kepler\ wavelength bands.
-We re-iterate that the extent to which nature produces starspot spectra resembling the prescription in Figure \ref{fig:filtercurve} remains an unknown, and is the main thrust of the proposed work effort.
-
-System-throughput-weighted contrast spectra can be integrated over the illustrated filter curves to yield precision estimates for contrasts in the two passbands.
-For this M0V example, we obtain $c_{\ktwo}\sim0.025$ and $c_{\tess}\sim0.1$.
-Starspots appear about four times darker in \kepler\ and \ktwo\ than they do in \tess.
-We define the ratio of starspot contrasts of \tess\ compared to \kepler\ as $\mathcal{R}\equiv \frac{c_{\tess}}{c_{\ktwo}}$.
-
-\noindent \textbf{Spot-induced modulation amplitudes depend on contrast.}
-
-Rotationally-modulated lightcurves exhibit semi-sinusoidal undulations as starspots (or spot groups) enter and exit the projected stellar disk.
-The most-spotted projected hemisphere is directed towards the observer at the moment of the lightcurve minimum, and the least-spotted projected hemisphere at lightcurve maximum.
-The amplitude of the lightcurve, therefore, is approximately the difference in starspot coverage fractions $\Delta f_{spot} \equiv f_{max}-f_{min}$ between the most-and-least spotted projected hemispheres, weighted by the spot contrast: $ A_\lambda = \Delta f_{spot} \cdot (1-c_\lambda)$.
-Therefore the ratio of \tess\ to \kepler\ amplitudes yields a constraint independent of the actual spot coverage fraction:
-
-$$ \frac{A_{TESS}}{A_{Kepler}} = \cancel{\frac{\Delta f_{spot}}{\Delta f_{spot}}} \cdot \frac{1-c_{TESS}}{1-c_{Kepler}} $$
-
-For our example with a 4500 K ambient photosphere and 2700 K starspot characteristic temperature, we obtained $c_{\ktwo}\sim0.025$ and $c_{\tess}\sim0.1$, yielding an expected amplitude ratio of:
-
-$$\frac{1-c_{TESS}}{1-c_{Kepler}} = \frac{1-0.1}{1-0.025}  = (0.9/0.975)=0.923$$
-
-\authorcomment1{We are not currently weighting by the TESS and K2 filter curves.  Here are the values we get when we take the mean:}
-
-$$\frac{1-c_{TESS}}{1-c_{Kepler}} = \frac{1-0.104}{1-0.00697}  = (0.896/0.99303)=0.902$$
-
-This equation is approximate, and in practice our team will derive $\mathcal{R}= \frac{c_{\tess}}{c_{\ktwo}}$ with Monte Carlo methods to take into account the uncertainty of the amplitude determination and subtle starspot self-dilution effects.
-\newline
 

### 2022-08-05 papers/paper1/ms.tex
Added equation
Uncomitted changes to the manuscript

@@
 Hypothetically, a contrast of unity would mean the starspot blends into ambient photosphere, its border vanishing entirely.
+
 Spot contrast is wavelength dependent ($c_\lambda$).
-It has a spectrum of its own, starting close to zero in the blue/visible, and approaching perhaps 40\% in the near-IR.
+There does not appear to be a first-principles theoretical prediction for the spectrum of a starspot as a function of the fundamental properties of the star.
+Arriving at such a theoretical relationship may be complicated.
+Instead, practitioners have come up with phenomenological models based on observed trends in spot contrast.
+Starspot flux is typically parameterized by labeling the emission components with a characteristic temperatures
+$T_\mathrm{spot}<T_\mathrm{phot}$.
+The spot spectrum depends primarily on that temperature: $S(T_\mathrm{spot})$, with additional minor correction terms for Zeeman broadening largely ignored for most applications possessing only low fidelity information of the spot spectrum.
+The wavelength-dependent spot contrast is then simply the ratio of the spot spectrum to the ambient photosphere spectrum:  $c_\lambda = \frac{S(T_\mathrm{spot})}{ S(T_\mathrm{phot})}$.
 
-There does not appear to be a first-principles theoretical prediction for the spectrum of a starspot as a function of the fundamental properties of the star.
-Instead, practitioners have come up with phenomenological models based on observed trends in spot contrast coupled with ostensibly commonsense heuristics.
-For example, most practitioners assume that starspot temperature depends principally on the photospheric temperature.
 Two sources provide similar but slightly different trends for this relationship.
@@
 Here we fit a quadratic relationship to the \citet{2016MNRAS.463.2494F} LAMOST-derived temperature points to obtain the relationship:
-\authorcomment1{Ryan H. to determine A, B, and C}.
 
 \begin{equation}
-  T_{\mathrm{spot}} = a \times T_{\mathrm{phot}}^2 + b \times T_{\mathrm{phot}} + c ~\mathrm{K}
+  T_{\mathrm{spot}} = 0.000117 \times T_{\mathrm{phot}}^2 + 1.37 \times T_{\mathrm{phot}} - 568 ~\mathrm{K}
   \label{eq:TspotFang}
@@
 
-We plot both trends in Figure \censorbox{X}.  The range of validity and accuracy of these relationships remain unclear. Interrogating these trends forms the main thrust of this work.
-
-Starspot flux can be parameterized by labeling the emission components with a characteristic temperatures
-$T_\mathrm{spot}<T_\mathrm{phot}$.The spot and ambient photosphere spectra depend only on those temperatures: $S(T_\mathrm{spot})$ and $S(T_\mathrm{phot})$.
-The wavelength-dependent spot contrast is then simply the ratio of the spot spectrum to the ambient photosphere spectrum:  $c_\lambda = \frac{S(T_\mathrm{spot})}{ S(T_\mathrm{phot})}$.
+The range of validity and accuracy of these relationships remain unclear.
 

### 2023-02-10 papers/paper1/ms.tex
Added citation for Cao

@@
 
-\authorcomment1{General background information and motivation for starspots...}
+Imperfections on the Sun's surface have been observed with telescopes for over 400 years \citep{1613idim.book.....G}.
+The mere existence of these sunspots betrayed an inexact understanding of the Sun, departing from the Platonic solid yellow circular disk that still prevails in cartoon pictures today.
+The existence of starspots on other stars was realized 75 years ago \citep{1947PASP...59..261K}, and has since enabled the entirely new subfields of gyrochronology, Doppler imaging, and stellar activity.
+Our community now takes for granted the omnipresence of starspots to measure rotation periods for 34,030 stars in the \emph{Kepler} mission \citep{2014ApJS..211...24M}, and about 29,860 stars in the \emph{K2} mission \citep{2020A&A...635A..43R}.
+The \emph{TESS} mission continues to collect lightcurves for millions of stars, making it destined to measure periods for hundreds of thousands of stars \citep{2022ApJ...927..219C}.
 
@@
 
-\authorcomment1{Todo: Cite Lyra Cao et al. in prep-- estimated August 2022}.
+In \citet{2022arXiv220910549C}, they borrow an equation from \citet{somers151}.
+In this equation, they define the variable $x_\mathrm{spot}$ as the ratio of $T_\mathrm{spot}$ over $T_\mathrm{ambient}$.
 

### 2023-04-29 papers/paper1/ms.tex
Updated environment, fleck, and tex files
Added Fleck to citations and tested fleck in the notebook

@@
 
-\begin{figure}[!htb]
+\begin{figure*}[!htb]
   \centering
-  \includegraphics[width=0.95\columnwidth]{Amplitude vs. Rotation for Kepler and TESS.png}
+  \includegraphics[width=0.95\textwidth]{Amplitude vs. Rotation for Kepler and TESS.png}
   \caption{Effect size for starspot contrast as seen in modulation amplitude versus period ($P_{\mathrm{rot}}$ [days]).
@@
 The blue and green dashed lines represents the average amplitudes for all plotted stars for both Kepler and TESS, approximately 3.98\% and 6.94\%, respectively.}
-\end{figure}
+\end{figure*}
 \FloatBarrier
@@
 \software{
-  pandas \citep{mckinney10, reback2020pandas},
-  emcee \citep{foreman13},
-  matplotlib \citep{hunter07},
   astroplan \citep{astroplan2018},
   astropy \citep{exoplanet:astropy13,exoplanet:astropy18},
+  bokeh \citep{bokehcite},
+  emcee \citep{foreman13},
   exoplanet \citep{exoplanet:exoplanet},
+  fleck \citep{fleckcite}
+  ipython \citep{perez07},
+  matplotlib \citep{hunter07},
   numpy \citep{harris2020array},
+  pandas \citep{mckinney10, reback2020pandas},
   scipy \citep{jones01},
-  ipython \citep{perez07},
-  bokeh \citep{bokehcite},
-  seaborn \citep{waskom14}}
+  seaborn \citep{waskom14}
+}
 %pytorch \citep{NEURIPS2019_9015}} % No pytorch yet!
